### 2016-10-28 paper.tex
first commit

%%
%% If you intend to use figures of formats jpg, png or pdf and want the
%% output to be immediately a pdf file, compile with pdflatex.
%%
%% If you want to use eps or ps figures, your output will be a dvi
%% file that can be converted to ps and pdf formats. In this case you
%% should compile your document with latex.
%%
%%This template is compatible with both methods.
%%

\documentclass[a4paper]{IEEEtran}
\usepackage{pdfpages}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[caption=false, ...]{subfig}

\title{{ \normalsize Master Thesis} \\
	Bluetooth Low Energy Supported Indoor Location}
\author{
	Ricardo Martins {\tt ricardo.pachecomartins@gmail.com}
	Instituto Superior T\'{e}cnico}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
 We are going to analyze the evolution of ethernet through the years up until the carrier ethernet. In order to do so we start from the very beginning of ethernet by presenting the bases of ethernet frames standardized in IEEE802.3 and from there on out we focus more on all the standards that were published in order to improve Virtual Local Area Networks (VLANs). We will show all the progress made in terms of scalability starting in small company VLANs and moving up until the most recent state of carrier ethernet that is capable of dealing with the needs of metro networks.

\end{abstract}

\section{Introduction}
\label{sec:introduction}
Neste paper iremos apresentar todo o estudo que foi feito para mostrar todo o trabalho que começou com a ethernet nos anos 70 e que progrediu até à carrier ethernet.

Para tal começamos por introduzir o que é a ethernet, obtendo assim conhecimentos sobre a base de estudo deste paper e continuamos de seguida apresentando o conceito de virtual local area networks. Voltando à ethernet, seguimos mostrando as evoluções que nela foram feitos para suportar a existência e melhor manipular VLANs, começando nos standards que permitiram suportar as VLANs de pequena dimensão até chegar ao ponto da carrier ethernet que permite a utilização desta tecnologia nas core networks.

Para melhor visualizar a evolução da ethernet até à carrier ethernet iremos olhar para diversos exemplos de como cada um dos standards apresentados se comporta com uma network exemplo.
 Desta forma será mais compreensível a visualização das suas variações em termos de funcionalidades e iremos também visualizar as contrapartidas de cada um dos standards em função da sua eficiência.

\section{Ethernet}
\label{sec:Ethernet}

Em 1973 Xerox começou a desenvolver uma rede local de computadores (LAN, Local Area Network) de topologia bus e em 1976 quando foi adicionado carrier sense, foi possível obter uma rede capaz de ligar mais de cem estações de trabalho com um cabo de um quilometro. Esta rede foi chamada de Ethernet e ganhou o seu nome da palavra “lumeniferous ether”\cite{ether} usada para descrever um meio passivo que transportaria ondas electromagnéticas no espaço, particularmente a luz do sol até à Terra.

Em 1980 foi criado um standard para a Ethernet de 10Mbps e a sua segunda versão foi apelidada de Ethernet II em 1982 que serviu de base para o IEEE802.3.

A ethernet  tem como objectivo interligar diferentes equipamentos no âmbito das LANs e
inclui a camada física e a camada de ligação de dados do modelo Open Systems Interconnection model (OSI model) . A camada de ligação de dados subdivide-se em duas partes, o Media Access Control(MAC) e o Logical Link Control(LLC).

A sub-camada MAC é responsável pelo controlo de acesso ao meio entre várias estações, permite receber e transmitir tramas. Cada estação tem um endereço MAC associado, um valor de 48 bits único que é atribuído pelo fabricante, que permite identificar a fonte e o destinatário das tramas. Através de códigos de redundância cíclica (CRC) consegue identificar erros e tentativas de manipulação nas tramas, sendo que as identificadas serão de seguida descartadas.

A sub-camada LLC  é responsável pela interface com as camadas superiores e pelo controlo de fluxo e de erros.

\section{Trama Ethernet}
\label{sec:Trama Ethernet}

A formata da trama Ethernet foi introduzido pelo IEEE802.3 já mencionado e apesar de todas as evoluções que foram inventadas para a Ethernet, o seu formatado permanece inalterado.

Composta por diversos parâmetros tem um tamanho variável entre os 84 bytes e os 1538 bytes como é possível verificar pela imagem~\ref{fig:EtherFrame}.

\textbf{Preâmbulo:} sequência de 7 bytes que permite a sincronização do receptor, pois este opera em modo burst.

\textbf{Start of Frame (SFD):}  Padrão de 8 bits (10101011) que indica o início da trama.

\textbf{Endereço de destino e endereço de fonte:} Endereços MAC, ambos com comprimento de 6 bytes.

\textbf{Comprimento/Tipo:}  sequência de 2 bytes que indica o comprimento, ou o tipo do campo de dados. Estes 2 bytes permitem a representação em base 10 dos valores 0 a 65535 e é em função deste valor que o significado deste campo se altera. Caso esteja entre 0 e 1500, então o  valor representa o comprimento do campo de dados e caso o valor seja maior que 1536 este campo apresenta o tipo da trama.

\textbf{Dados:} Campo reservado para guardar toda a informação útil a ser transmitida. Este campo tem como tamanho mínimo 46 bytes podendo chegar até aos 1500. Na realidade o tamanho mínimo de informação que pode ser transmitido é de 40 bytes, sendo que os restantes 6 bytes necessários para obter o valor mínimo deste campo serão preenchidos com padding a zero.

\textbf{Frame Check Sequence (FCS)}: Conjunto de 4 bytes que contem o valor do CRC calculado sobre todos os bits da trama com excepção do preâmbulo e SFD. O CRC tem como objetivo permitir ao receptor garantir que a informação da trama não foi alterada durante a sua transmissão.

A trama apresentada tem um comprimento que difere em 12 bytes do valor apresentado no início do capítulo, esta diferença deve-se ao facto que um switch tem que esperar uma quantidade igual ao período de transmissão de 96 bits (12 bytes). Assim sendo este período de tempo depende da velocidade de transmissão.

\section{Virtual LAN}
\label{sec:Virtual LAN}

 \begin{figure*}
 	\centering
 	\includegraphics[width=0.75\textwidth,height=6cm]{./VLAN_ex}
 	\caption{Exemplo de rede para IEEE802.1Q}
 	\label{fig:VLANex}
 \end{figure*}

Uma LAN virtual (VLAN) consiste no agrupamento lógico de vários nós (dispositivos de rede), sobre uma determinada topologia física, de modo a criar a nível lógico uma nova rede com todas as funcionalidades de uma LAN. Desta forma os administradores de rede têm a capacidade de criar divisórias nas suas redes que concretizem as necessidades funcionais e de segurança sem haver necessidade de adicionar mais cabos ou efectuar mudanças na infraestructura da rede.
A criação de VLANs traz consigo um número de vantagens como por exemplo:

\textbf{Broadcast Control:} Broadcasts são necessários para o normal funcionamento de uma rede e muitos protocolos e aplicações depende deste mecanismo para funcionar. Para redes de grande dimensão este processo pode consumir muitos recursos. Ao segmentar uma LAN de grande dimensão em pequenas VLANs é possível reduzir o tráfico de broadcast visto que cada uma das mensagens de broadcast passam a ser enviadas apenas para os membros da VLAN na qual se encontra quem a enviou. Cada grupo passa a ter a sua própria spanning tree e define um domínio de difusão.

\textbf{Segurança :} VLANs permitem aumentar a segurança e privacidade da rede ao separar o tráfego pertencente a diferentes organizações/departamentos. Num ambiente de rede VLAN, com múltiplos broadcast domains, os administradores de redes têm a capacidade de controlar cada porto e utilizador. Desta maneira deixa de ser possível que um utilizador malicioso apenas se ligue a um porto de switch para visualizar o tráfego da rede.

\textbf{Transparência da camada física :} Facilita a gestão da rede na medida que permite ao administrador da rede organizar os utilizadores em grupos de modo a reflectir a estrutura da organização (departamentos, edifícios, etc.), de modo independente da topologia física

\textbf{Custo: }Segmentar um VLAN de grande dimensão em VLANs de menor dimensão é mais barato que criar uma rede com routers, visto que caso geral, routers são mais caros que switches

Uma VLAN pode ser de três tipos diferentes:

\textbf{Port-based VLAN:} Uma VLAN fica associada a um porto de um switch. Desta forma cada estação ligada a este porto passa automaticamente a pertence à VLAN associada.

\textbf{MAC Address-based VLAN:} As VLANs agregam as estações de acordo com o endereço MAC. Esta relação é introduzida via software pelo administrador de rede. Desta forma um estação pode pertencer a diferentes VLANs.

\textbf{Network Address-based VLAN:} O agrupamento de estações numa VLAN é realizado através da informação contida no endereço do cabeçalho da camada de rede. Para este tipo de VLANs os switches têm que ser capazes de processar os cabeçalhos da camada 3, ou apenas opta-se por utilizar routers.

Para garantir o funcionamento de uma rede com switches, cada um deles tem que ser capaz de localizar as estações através da tabela de encaminhamento e de conhecer a VLAN a que pertencem o destinatário e a fonte de cada mensagem. Para redes de grande dimensão as tabelas de encaminhamento de cada switch tornam-se problemáticas, piorando o desempenho da rede. 
Houve assim necessidade de etiquetar as tramas de acordo com a VLAN, solução apresentada com o standard IEEE802.1Q, descrito no capitulo~\ref{sec:802.1Q}.

\section{Standard IEEE802.1Q}
\label{sec:802.1Q}

O standard mais simples para fornecer uma Ethernet Virtual Private LAN (E-LAN), “Multipoint-to-Multipoint”, é o IEEE802.1Q. Este standard permite criar VLANs numa estructura LAN, permitindo empresas controlar e separar trafego entre diferentes departamentos.
Para obter estas funcionalidades é necessário adicionar um parâmetro às tramas utilizadas (visualizadas no capitulo ~\ref{sec:Trama Ethernet}), uma etiqueta (tag) de 4 octetos, denominada Q-tag.

Para tal é necessário adicionar uma etiqueta (tag) de 4 octetos , designada por etiqueta - Q (Q tag), a seguir ao campo de endereço fonte da trama 802.3. Vamos então analisar a trama utilizada neste standard, apresentada na imagem ~\ref{fig:TramaQ}.

\begin{figure}[htp]
	\centering
	\includegraphics[width=1\columnwidth]{./802_1Q}
	\caption{Trama utilizada no IEEE802.1Q}
	\label{fig:TramaQ}
\end{figure}

\textbf{Identificador do protocolo VLAN:} Valor de 16 bits utilizado para identificar o protocolo VLAN. Para  IEEE802.1Q trama com Q-tag o valor é 0x8100.

\textbf{Prioridade (PCP, Priority control point):} Campo de 3 bits que definem o nível de prioridade da trama. Este valor é utilizado para dar prioridades diferentes a certo tipos de data (voice,video,etc.)

\textbf{Identificador de formato canónico (CFI) :} Campo de 1 bit utilizado para o propósito de compatibilidade entre ethernet (valor 0)  e token ring (valor 1). Este campo já não é mais utilizado.

\textbf{Identificador VLAN:} campo de 12 bits que especifica a VLAN à qual a trama pertence.

Este standard está limitado em termos de escalabilidade pelo parâmetro identificador VLAN, visto que é composto somente por 12 bits. Assim sendo só é possível ter 4094 VLAN diferentes pois os valores 0x000, que indica que um frame não tem VLAN ID, e 0xFFF, reservado para implementação, não podem ser utilizados.
Devido a esta limitação este standard pode funcionar nas restrições de uma só organização, mas rapidamente se torna obsoleto quando os “Service Providers” têm como objetivo distribuir serviços ethernet para múltiplos end-users numa infrastructura de rede partilhada.

Tendo analisado o standard IEEE802.1Q podemos então analisar o seu funcionamento em função da imagem ~\ref{fig:VLANex}.

 Tendo analisado o standard podemos então visualizar o seu funcionamento através da figura x.

 Estamos então perante um cenário onde há três VLANs diferentes, distribuídas aleatoriamente na rede. Para cada uma das VLANs há um exemplo de transmissão de informação, representado com as setas de cores, que demonstra como é feita a comunicação entre estações na mesma VLAN.

 Vamos então seguir o exemplo de uma mensagem da VLAN1 representada pela cor verde. Neste caso uma estação na VLAN1 mais à esquerda decide enviar uma mensagem para toda as restantes estações da sua VLAN e para tal envia uma mensagem broadcast. A mensagem para as estações dentro da sua “nuvem” estão já recebidas e falta só chegar a todas as restantes na rede, para tal, a mensagem segue até ao seu switch associado S1 e este irá reencaminhar a mensagem para o resto da rede. O standard IEEE802.1Q garantiu que será o switch S1 a realizar o Q-tagging, isto é, inserir a tag 1 (cor verde) na trama enviada, como pode ser visto na representação da trama na figura X. A trama é encaminhada para o switch principal que segundo a sua tabela de encaminhamento irá transmitir a trama para todos os switches que tenham a VLAN1 nos seus portos, ou seja os switches S3 e S4. Quando a mensagem chega aos switches S3 e S4, eles retiram a tag e reencaminham a mensagem para as estações da VLAN1 que eles conhecem.

 \begin{figure*}
 	\centering
 	\includegraphics[width=0.75\textwidth,height=6cm]{./ad_ex}
 	\caption{Exemplo de rede para IEEE802.1ad}
 	\label{fig:adex}
 \end{figure*}

\section{Standard IEEE802.1ad}
\label{sec:802.1ad}

Devido às limitações nas capacidades de escalonamento criou-se o standard IEEE802.1ad, também conhecido como "QinQ". Para analisar as inovações apresentadas em comparação com o seu standard antecessor, IEEE802.1Q, basta olhar para a figura ~\ref{fig:trama_ad}.

Como é possível visualizar na figura, a grande diferença entre os standards foi a adição de uma segunda “Q-TAG” (campo apresentado no capitulo ~\ref{sec:802.1Q} com o IEEE802.1Q) entre o customer tag e o endereço de origem e a distinção entre os dois agora existentes. O antigo campo de 4 bytes ficou conhecido como o Customer Tag e o novo campo como Service Tag.

\begin{figure}[htp]
	\centering
	\includegraphics[width=1\columnwidth]{./trama_ad}
	\caption{Trama utilizada no IEEE802.1ad}
	\label{fig:trama_ad}
\end{figure}

Este ID, Service tag, é utilizado para identificar o serviço na rede fornecedora, enquanto que o VLAN ID do cliente, designado como C-tag, permanece intacto e não é alterado pelo fornecedor do serviço. 
Isto permite ao C-tag ser transparente na rede. Os Provider Bridges utilizam o S-tag para identificar a que serviço as frames de um cliente pertencem, logo, cada serviço requer um S-tag distinto. Este standard foi criado por algumas razões importantes:

\textbf{Aumento do limite de VLANs}: Este standard apresenta uma frame com duas tags passando a ter uma limitação teórica de 4096x4096=16777216 LAN’s, permitindo um acomodamento maior da rede.

\textbf{Compatibilidade:} compatível com o seu antecessor (802.1q) e, embora seja limitado a duas tags, não existe um tecto máximo no standard que limite um frame a mais do que duas tags, permitindo um crescimento no protocolo.É mais fácil para os fabricantes de equipamento de redes modificar o equipamento já existente criando múltiplos cabeçalhos 802.1q do que modificar o equipamento para implementar alguma hipotética extensão do cabeçalho da VLAN ID que não seja do formato 802.1q.

Contudo este standard não corrigiu todas as limitações existentes apresentadas no capitulo XX.

Continua a não haver separação entre as MAC addresses do fornecedor e do cliente, o que leva a que possa haver um problema de segurança, uma vez que a informação do seu endereçamento está visível fora do domínio de segurança da rede. 
Outra das limitações é que cada switch tem de saber todos os “MAC addresses”, o que implica que cada switch Provider Bridge deverá ter capacidade para suportar centenas de milhares de endereços, o que vai introduzir problemas de escalabilidade. Um exemplo representativo deste problema é no caso em que queremos adicionar um novo utilizador à rede, sendo que tem que ser memorizado pelos switches do fornecedor caso contrário, quando ocorrer uma falha na rede do consumidor, a ação tomada pelo Spanning Tree Protocol (STP) pode afectar a rede do consumidor. Os "downtimes" da rede em caso de falham continuam dependentes da execução completa da STP, mantendo assim as limitações que não permitem que a ethernet seja utilizada nas redes metropolitanas.

Podemos então passar a visualizar a figura ~\ref{fig:adex} de modo a perceber em termos funcionais o que foi alterado. Começamos por denotar que existem dois tipos diferentes de switches, os customer bridges (CB), switches que utilizam o IEEE802.1Q e os Provider bridges (PB), que funcionam com o IEEE802.1ad agora exemplificado.

Um edge switch, nome atribuído aos switches que estão na fronteira, isto é, comunicam com customer bridges, tem a funcionalidade de inserir e remover a service tag das tramas que por eles passam, ou seja, ao receber uma trama vinda de um CB é inserida uma S-tag correspondente ao serviço associado à trama. De seguida a trama será reencaminhada dentro desta nuvem que engloba os PBs todos, chegando a um switch VLAN-aware. Esta característica usada para descrever certos switches, indicam que o switch em causa conhece toda a rede e é capaz de reencaminhar correctamente a trama para o seu destino. Ao chegar de novo a um edge switch, o mesmo irá retirar a S-tag da trama e reencaminhar para o CB adjacente.

\section{Standard IEEE802.1ah}
\label{sec:802.1ah}

O IEEE802.1ah, denominado de Provider Backbone Bridges (PBB) ou MAC-in-MAC, sucessor do standard apresentado no último capitulo evoluiu a ethernet ao adicionar endereços MAC dedicados para os service providers.  Pelo figura ~\ref{fig:trama_ah} é possível visualisar todas as alterações que foram feitas com este standard:

\begin{figure}[htp]
	\centering
	\includegraphics[width=1\columnwidth]{./ah_trama}
	\caption{Trama utilizada no IEEE802.1ah}
	\label{fig:trama_ah}
\end{figure}

\textbf{endereço de destino backbone}: campo de 6 bytes que contem o endereço MAC do switch PBB destino;

\textbf{endereço de fonte backbone}: campo de 6 bytes que contem o endereço MAC do switch PBB fonte;

\textbf{identificador de VLAN backbone( B-VID):} campo de 2 bytes utilizado para identificar o indentificador de VLAN atribuido à trama

\textbf{identificador de serviço:} campo de 3 bytes utilizador para indicar o serviço da trama

Todas estas adições no formato da trama permitiram ao IEEE802.1ah:

Oferecer a completa separação entre os MAC addresses do fornecedor e do cliente;
Não impõr qualquer alteração ao processo de switching da Ethernet nas pontes fundamentais;
Proporcionar uma clara separação entre o domínio do consumidor e do fornecedor;
Deduzir os MAC addresses do consumidor apenas através dos Backbone Edge Bridges;
Suportar até 16 milhões de VLAN’s (campo service-id de 24 bits).

Contudo, este standard também tem desvantagens, e uma das mais evidentes é o facto das redes PBB continuarem a utilizar o Spanning Tree Protocol para eliminar malhas fechadas.

\section{Standard IEEE802.1ag}
\label{sec:802.1ag}

o standard IEEE802.1ag, uma actualização significativa do IEEE802.1ah, será o último a ser apresentado pois foi com ele que se obteve a chamada carrier ethernet. Baseado na trama utilizada pelo IEEE802.1ah e sem necessidade de adição de novos campos apresentou um número elevado de inovações que permitiu finalmente que a ethernet fosse utilizada como uma opção viável nas redes metropolitanas.

\textbf{Engenharia de tráfego:} tomando vantagem do facto que ao desligar algumas das funcionalidades, o hardware existente ethernet é capaz de ter um novo comportamento de reencaminhamento, ou seja, é capaz de realizar um reencaminhamento de ligação orientada.

Os ethernet switches baseiam-se nos 60 bits , endereço MAC (48 bits) mais o identificador VLAN (12bits), para reencaminhar mensagens, sendo que os dois parâmetros são únicos globalmente. Visto que  o identificador VLAN é utilizado para identificar um domínio sem loops no qual os endereços MAC são "flooded", se optarmos por configurar rotas sem loops com endereços MAC, o identificador VLAN é libertado das suas funções. Assim sendo no IEEE802.1ag este parâmetro é utilizado para identificar rotas específicas na rede para um dado endereço MAC destino.

\begin{figure}[htp]
	\centering
	\includegraphics[width=1\columnwidth]{./te_ag}
	\caption{Visualização da engenharia de tráfego}
	\label{fig:te_ag}
\end{figure}

Este standard permite que certos endereços tenham as suas tabelas de encaminhamento preenchidas pelo plano de controlo em vez de ser pela utilização do STP. O funcionamento de cada switch mantem-se intacto, a diferença é que já não há necessidade de mecanismos de aprendizagem, resultando em rotas predeterminadas pela rede e um comportamento totalmente previsível da rede sobre qualquer circunstância.
A figura ~\ref{fig:te_ag} permite visualizar a engenharia de tráfego aplicada à rede no sentido em que é visível duas rotas para o mesmo endereço MAC de destino, o working path que representa o caminho habitualmente utilizado e o Protection Path (indicado por um identificador VLAN diferente) que é utilizado em caso de falha da rota principal. Esta implementação permitiu reduzir drasticamente o tempo de baixa da rede para situações de falha de rota, visto que ao invés de ser necessário esperar pelo STP bastanta reencaminhar o tráfego pelo working path.

\textbf{Operações, administração e gestão (OAM):} Funcionalidades OAM estão bem definidas nas redes TDM e são um factor importante para garantir que as operadoras pode oferecer serviços de qualidade "carrier". Os OAMs estão divididos em duas áreas, gestão de falha e monitorização de performance ,vamos então ver o que o IEEE802.1ag melhorou nestes campos:

\textbf{Gestão de falha:}Garante que quando ocorre uma falha na rede, o problema é encaminhado para o operador que depois toma a decisão mais correcta. Está divida em três partes:

\textbf{deteção de falha:} Mensagens de verificação contínua (continuous check messages (CCMs)) são enviadas pela fonte para o destino em intervalos fixos de tempos. Se uma resposta não for recebido dentro de um período específico de tempo (normalmente 3 vezes o intervalo de envio), anuncia-se uma falha na rede. O limite de tempo imposto para recuperação da rede nas redes metropolitanas é de 50ms, este valor era impossível de obter com o STP. Com as CCMs ao definir o intervalo de envio em valores inferior ao limite, é possível detectar uma falha em menos de 50ms.

\textbf{verificação da falha:} Mensagens de loopback (LBM) e respostas loopback (LBR) permitem descobrir se uma determinada rota funciona. Normalmente utilizadas após a criação de uma rota ou quando se detecta um falha, para verificar que ela ocorre entre dois endpoints.

\textbf{isolação de falha:} Mensagens de link trace (LTM) e respostas link trace (LTR) permitem descobrir qual a ligação de uma certa rota que falhou. Ao enviar esta mensagem (LBM) numa determinada rota que contem uma falha, todos os switches que a receberem devem responder com uma LBR, deste modo é possível identificar a partir de que switch a mensagem parou de ser encaminhada, identificando a ligação partida.

\textbf{Monitorização de performance:} As tramas têm diferentes níveis de performance visto que num serviço podem sofrer de delays devido a congestionamento, ou mesmo perdas de tramas. Serviços de vídeo e de voz são os mais susceptíveis  aos efeitos de latência e jitter. Assim sendo as redes de carrier ethernet requerem monotorização de performance avançadas de modo a ajudar os provedores de serviço a cumprir os requerimentos da rede.

\textbf{rácio de tramas perdidas:} Conhecido como Frame loss ratio, é calculado usando CCMs e contadores. Cada um dos endpoints envia um certo número de CCMs e conta quantas é que foram recebidas. No fim os contadores são trocados para calcular a percentagem de tramas perdidas.

\textbf{atrasos nas tramas:} Mensagens são enviadas com os tempos de envio. O switch destino pode então calcular o tempo que a trama demorou até ser recebida. Só funciona se ambos os switches tiverem os relógios sincronizados.

\textbf{ variação dos atrasos nas tramas:} Mesmo conceito que o apresentado anteriormente excepto que são armazenados os valores dos delays de modo a ver a sua variação.

\section{Comparação de eficiência}
\label{sec:eficiencia}

Até este ponto foram analisados todos os standard criados para a obtenção da carrier ethernet e os benefícios que apresentaram sem nunca ter focado nas contra-partidas. Com cada um destes standards houve necessidade de inserir mais informação na trama, aumentando a carga que tem que ser transportada com cada mensagem. Iremos então analisar os efeitos de cada standard na eficiência.

A eficiência de um standard depende da seguinte fórmula:

\begin{equation}
	\label{eq:problem}
	 \text {Eficiencia} = \frac{\text{tamanho  da informação útil}}{ \text{tamanho da trama}} \\ 
\end{equation}\\

Relembramos que a informação tem um valor variável entre 40 e 1500 bytes , como visto na secção XXX e que o tamanho da trama varia de standard para standard.
Como ponto de partida para todas estas inovações estava a ethernet e portanto iremos analisar a sua eficiência como base de comparação.

A trama ethernet apresentada na figura ~\ref{fig:EtherFrame} tem necessidade de 38 bytes de overhead que são sempre adicionados à data. Assim sendo podemos visualizar a variação da sua eficiência na figure ~\ref{fig:ef_802.3}.

\begin{figure}[htp]
	\centering
	\includegraphics[width=1\columnwidth]{./ef_802_3}
	\caption{ eficiência do IEEE802.1Q}
	\label{fig:ef_802.3}
\end{figure}

Para os standards analisados neste paper, as suas eficiências são comparadas na figura ~\ref{fig:ef_all}.
O standard IEEE802.1Q está representado com a cor azul e é o mais eficiente em comparação aos restantes devido a só ter sido adicionado um campo de 4 bytes na trama, totalizando a um overhead de 42 bytes. De seguida está o IEEE802.1ad com a cor vermelha, standard que adicionou um segundo campo de 4 bytes, para um overhead de tamanho 46 bytes. Por último está o IEEE802.1ah com a cor verde, que apresenta a pior eficiência de todos os standards apresentados devido ao tamanho total do overhead de 67 bytes que impõe às suas tramas.

\begin{figure}[htp]
	\centering
	\includegraphics[width=1\columnwidth]{./ef_all}
	\caption{eficiência do IEEE802.1Q,IEEE802.1ad e IEEE802.1ah}
	\label{fig:ef_all}
\end{figure}

\section{Conclusions}
\label{sec:conclusion}

A ethernet como foi estudada no capítulo ~\ref{sec:Ethernet} tinha vários defeitos que a impediam de ser implementada nas redes metropolitanas. Através de todas as inovações que foram implementas, chegando ao nível apresentado no IEEE802.1ag, já é possível. De momento a ethernet já não está limitada pelos seus problemas de escalabilidades, apresenta altos níveis de segurança e permite obter um rede resiliente capaz de monotorização de performance e gestão de falhas, cumprindo assim os requisitos em termos de OAMs

\begin{thebibliography}{9}

\bibitem{ether}
 \texttt{http://listserv.linguistlist.org/pipermail/ads-l/1999-July/001027.html}
\bibitem{Carrier}
 \texttt{Ethernet as a Carrier Grade Technology: Developments and Innovations} Rafael Sánchez,2008
 \bibitem{slides}
 \texttt{Slides da unidade curricular- Redes de Telecomunicações- Capitulo 3} João Pires

\end{thebibliography}

\end{document}
This is never printed

### 2016-10-28 paper.tex
Preparing bullet points

@@
 \section{Introduction}
-\label{sec:introduction}
-Neste paper iremos apresentar todo o estudo que foi feito para mostrar todo o trabalho que começou com a ethernet nos anos 70 e que progrediu até à carrier ethernet.
+\label{sec:Introduction}
+Presenting the problem:
+	- From GPS to Indoor Location Problems
+	- Various Solutions
+	- BLE Problems for Indoor Location
+	- Lack of FrameWork
+	- Presenting the papers structure
 
-Para tal começamos por introduzir o que é a ethernet, obtendo assim conhecimentos sobre a base de estudo deste paper e continuamos de seguida apresentando o conceito de virtual local area networks.
-Voltando à ethernet, seguimos mostrando as evoluções que nela foram feitos para suportar a existência e melhor manipular VLANs, começando nos standards que permitiram suportar as VLANs de pequena dimensão até chegar ao ponto da carrier ethernet que permite a utilização desta tecnologia nas core networks.
-
-Para melhor visualizar a evolução da ethernet até à carrier ethernet iremos olhar para diversos exemplos de como cada um dos standards apresentados se comporta com uma network exemplo.
- Desta forma será mais compreensível a visualização das suas variações em termos de funcionalidades e iremos também visualizar as contrapartidas de cada um dos standards em função da sua eficiência.
-
-\section{Ethernet}
-\label{sec:Ethernet}
-
-Em 1973 Xerox começou a desenvolver uma rede local de computadores (LAN, Local Area Network) de topologia bus e em 1976 quando foi adicionado carrier sense, foi possível obter uma rede capaz de ligar mais de cem estações de trabalho com um cabo de um quilometro.
-Esta rede foi chamada de Ethernet e ganhou o seu nome da palavra “lumeniferous ether”\cite{ether} usada para descrever um meio passivo que transportaria ondas electromagnéticas no espaço, particularmente a luz do sol até à Terra.
-
-Em 1980 foi criado um standard para a Ethernet de 10Mbps e a sua segunda versão foi apelidada de Ethernet II em 1982 que serviu de base para o IEEE802.3.
-
-A ethernet  tem como objectivo interligar diferentes equipamentos no âmbito das LANs e
-inclui a camada física e a camada de ligação de dados do modelo Open Systems Interconnection model (OSI model) .
-A camada de ligação de dados subdivide-se em duas partes, o Media Access Control(MAC) e o Logical Link Control(LLC).
-
-A sub-camada MAC é responsável pelo controlo de acesso ao meio entre várias estações, permite receber e transmitir tramas.
-Cada estação tem um endereço MAC associado, um valor de 48 bits único que é atribuído pelo fabricante, que permite identificar a fonte e o destinatário das tramas.
-Através de códigos de redundância cíclica (CRC) consegue identificar erros e tentativas de manipulação nas tramas, sendo que as identificadas serão de seguida descartadas.
-
-A sub-camada LLC  é responsável pela interface com as camadas superiores e pelo controlo de fluxo e de erros.
+\section{Related Work}
+\label{sec:RelatedWork}
+	- Brief BLE Presentation;
+	- BLE Indoor Solution and Existing Applications
+	-
 

### 2016-11-10 paper.tex
intro first draft

@@
 \begin{abstract}
- We are going to analyze the evolution of ethernet through the years up until the carrier ethernet.
-In order to do so we start from the very beginning of ethernet by presenting the bases of ethernet frames standardized in IEEE802.3 and from there on out we focus more on all the standards that were published in order to improve Virtual Local Area Networks (VLANs).
-We will show all the progress made in terms of scalability starting in small company VLANs and moving up until the most recent state of carrier ethernet that is capable of dealing with the needs of metro networks.
+
+	Abstract
 
@@
 \section{Introduction}
-\label{sec:Introduction}
-Presenting the problem:
-	- From GPS to Indoor Location Problems
-	- Various Solutions
-	- BLE Problems for Indoor Location
-	- Lack of FrameWork
-	- Presenting the papers structure
+\label{sec:intro}
 
-\section{Related Work}
-\label{sec:RelatedWork}
-	- Brief BLE Presentation;
-	- BLE Indoor Solution and Existing Applications
-	-
+Indoor positioning systems have greatly evolved in the past few years due to the great success of its counter-part, the global positioning system (GPS), in outdoor environments but failure to reproduce the same results in indoor environments.
+Since GPS is an outdoor position systems and is based on a network of satelite, when the required scenario for position tracking is inside a building, new constraints are presented onto the process such as the attenuation and reflection of eletromagnetic waves upon collision with building walls and obstacles \cite{surveygps}.
+As such there was a need to find reliable indoor systems that by nature would already be able to heavily reduce the impact of some of the mentioned constraints.
 
-\section{Trama Ethernet}
-\label{sec:Trama Ethernet}
+In order to understand indoor position there is a need to understand the full scope of variables that come to surface when moving from outdoor to indoor.
+When developing a system there is a need to make sure that it can tackle the chalenges such as small space dimension which reforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques}.
+All of the previously mentioned affect the way electromagnetic waves propagate in an indoor environment leading to problems related to severe multipath and reflection on existent surfaces \cite{surveywireless}.
+Besides propagation challenges, there are energy consumption, accuracy and deployment costs that play a critical role in deciding the viability of a proposed indoor location technique.
 
-A formata da trama Ethernet foi introduzido pelo IEEE802.3 já mencionado e apesar de todas as evoluções que foram inventadas para a Ethernet, o seu formatado permanece inalterado.
+With the evolution of mobile devices there has been a sizeable number of different technologies that can possibly be used in indoor locationing \cite{surveywireless,survey2,survey1} such as GPS-based technologies, using high sensitivity antenas to overcome GPS's indoor issues, RFID , Wireless LAN and Bluetooth among others, allowing even for hibrid systems which make use of more than one of the technologies mentioned above.
 
-Composta por diversos parâmetros tem um tamanho variável entre os 84 bytes e os 1538 bytes como é possível verificar pela imagem~\ref{fig:EtherFrame}.
+Another important aspect besides the chosen technology is the location detection technique which are widely varied in terms of accuracy and complexity.
+The simplest available category would be proximity detection which has Cell of Origin (CoO) as its most famous technique, allowing for room-based detection by assuming that the mobile target is at the cell with the strongest receiving power.
+This technique is widely used by system using RFID, Bluetooth and Infrared.
+The next category would be triangulation which uses the geometric properties of the triangle to obtain a position.
+The method utilized to obtain the values used in the position's calculation can be angle-based, which have implementation costs are their worst limitation; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI) as a metric, altough it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position utilizing the last determined position and incrementing based on estimate of the device's speed.
 
-\textbf{Preâmbulo:} sequência de 7 bytes que permite a sincronização do receptor, pois este opera em modo burst.
+The solution in this paper is based on the Bluetooth Low energy technology using the CoO method to obtain the device's location and as such for what technologies and location methods are concerned, only the used ones will be focused.
 
-\textbf{Start of Frame (SFD):}  Padrão de 8 bits (10101011) que indica o início da trama.
+In this paper is structured in a way that in \ref{sec:related} an overview of Bluetooth low energy (BLE) and existent projects and their associated architecture are presented while shedding light on the benefits and limitations of the chosen technologies, in \ref{sec:structure} the proposed system is described based on technology and method used as well as the architecture of the would system, \ref{sec:performance} analyzes the different aspects of the presented solution in terms of energetic efficiency, accuracy and response timings, finalizing in \ref{sec:future} by overviewing what could be carried out in order to further develop the existing work and by concluding the paper.
 
-\textbf{Endereço de destino e endereço de fonte:} Endereços MAC, ambos com comprimento de 6 bytes.
+- Falta de framework, não tenho bases, vale a pena mencionar na introdução?
 
-\textbf{Comprimento/Tipo:}  sequência de 2 bytes que indica o comprimento, ou o tipo do campo de dados.
-Estes 2 bytes permitem a representação em base 10 dos valores 0 a 65535 e é em função deste valor que o significado deste campo se altera.
-Caso esteja entre 0 e 1500, então o  valor representa o comprimento do campo de dados e caso o valor seja maior que 1536 este campo apresenta o tipo da trama.
+\section{Related Word}
+\label{sec:related}
 
-\textbf{Dados:} Campo reservado para guardar toda a informação útil a ser transmitida.
-Este campo tem como tamanho mínimo 46 bytes podendo chegar até aos 1500.
-Na realidade o tamanho mínimo de informação que pode ser transmitido é de 40 bytes, sendo que os restantes 6 bytes necessários para obter o valor mínimo deste campo serão preenchidos com padding a zero.
+Bluetooth is a wireless technology that was created in 1994 with the objective of replacing cables connecting fixed or portable devices.
+At this point in time Bluetooth Special Interest Group is in charge of developing and managing this technology characterized by its robustness, low energy consuption and low cost.
+The Bluetooth Low Energy protocol was introduced with the Bluetooth Core Specification version 4 (also called Bluetooth Smart) circa 2010 alongside two other protocols.
+Out of the three, BLE standed out for its lower power consuption, lower complexity and lower cost, while allowing for  device discovery, connection establishment and connection mechanisms.
 
-\textbf{Frame Check Sequence (FCS)}: Conjunto de 4 bytes que contem o valor do CRC calculado sobre todos os bits da trama com excepção do preâmbulo e SFD.
-O CRC tem como objetivo permitir ao receptor garantir que a informação da trama não foi alterada durante a sua transmissão.
-
-A trama apresentada tem um comprimento que difere em 12 bytes do valor apresentado no início do capítulo, esta diferença deve-se ao facto que um switch tem que esperar uma quantidade igual ao período de transmissão de 96 bits (12 bytes).
-Assim sendo este período de tempo depende da velocidade de transmissão.
-
-\section{Virtual LAN}
-\label{sec:Virtual LAN}
-
- \begin{figure*}
- 	\centering
- 	\includegraphics[width=0.75\textwidth,height=6cm]{./VLAN_ex}
- 	\caption{Exemplo de rede para IEEE802.1Q}
- 	\label{fig:VLANex}
- \end{figure*}
-
-Uma LAN virtual (VLAN) consiste no agrupamento lógico de vários nós (dispositivos de rede), sobre uma determinada topologia física, de modo a criar a nível lógico uma nova rede com todas as funcionalidades de uma LAN.
-Desta forma os administradores de rede têm a capacidade de criar divisórias nas suas redes que concretizem as necessidades funcionais e de segurança sem haver necessidade de adicionar mais cabos ou efectuar mudanças na infraestructura da rede.
-A criação de VLANs traz consigo um número de vantagens como por exemplo:
-
-\textbf{Broadcast Control:} Broadcasts são necessários para o normal funcionamento de uma rede e muitos protocolos e aplicações depende deste mecanismo para funcionar.
-Para redes de grande dimensão este processo pode consumir muitos recursos.
-Ao segmentar uma LAN de grande dimensão em pequenas VLANs é possível reduzir o tráfico de broadcast visto que cada uma das mensagens de broadcast passam a ser enviadas apenas para os membros da VLAN na qual se encontra quem a enviou.
-Cada grupo passa a ter a sua própria spanning tree e define um domínio de difusão.
-
-\textbf{Segurança :} VLANs permitem aumentar a segurança e privacidade da rede ao separar o tráfego pertencente a diferentes organizações/departamentos.
-Num ambiente de rede VLAN, com múltiplos broadcast domains, os administradores de redes têm a capacidade de controlar cada porto e utilizador.
-Desta maneira deixa de ser possível que um utilizador malicioso apenas se ligue a um porto de switch para visualizar o tráfego da rede.
-
-\textbf{Transparência da camada física :} Facilita a gestão da rede na medida que permite ao administrador da rede organizar os utilizadores em grupos de modo a reflectir a estrutura da organização (departamentos, edifícios, etc.), de modo independente da topologia física
-
-\textbf{Custo: }Segmentar um VLAN de grande dimensão em VLANs de menor dimensão é mais barato que criar uma rede com routers, visto que caso geral, routers são mais caros que switches
-
-Uma VLAN pode ser de três tipos diferentes:
-
-\textbf{Port-based VLAN:} Uma VLAN fica associada a um porto de um switch. Desta forma cada estação ligada a este porto passa automaticamente a pertence à VLAN associada.
-
-\textbf{MAC Address-based VLAN:} As VLANs agregam as estações de acordo com o endereço MAC.
-Esta relação é introduzida via software pelo administrador de rede.
-Desta forma um estação pode pertencer a diferentes VLANs.
-
-\textbf{Network Address-based VLAN:} O agrupamento de estações numa VLAN é realizado através da informação contida no endereço do cabeçalho da camada de rede.
-Para este tipo de VLANs os switches têm que ser capazes de processar os cabeçalhos da camada 3, ou apenas opta-se por utilizar routers.
-
-Para garantir o funcionamento de uma rede com switches, cada um deles tem que ser capaz de localizar as estações através da tabela de encaminhamento e de conhecer a VLAN a que pertencem o destinatário e a fonte de cada mensagem.
-Para redes de grande dimensão as tabelas de encaminhamento de cada switch tornam-se problemáticas, piorando o desempenho da rede.
-Houve assim necessidade de etiquetar as tramas de acordo com a VLAN, solução apresentada com o standard IEEE802.1Q, descrito no capitulo~\ref{sec:802.1Q}.
-
-\section{Standard IEEE802.1Q}
-\label{sec:802.1Q}
-
-O standard mais simples para fornecer uma Ethernet Virtual Private LAN (E-LAN), “Multipoint-to-Multipoint”, é o IEEE802.1Q.
-Este standard permite criar VLANs numa estructura LAN, permitindo empresas controlar e separar trafego entre diferentes departamentos.
-Para obter estas funcionalidades é necessário adicionar um parâmetro às tramas utilizadas (visualizadas no capitulo ~\ref{sec:Trama Ethernet}), uma etiqueta (tag) de 4 octetos, denominada Q-tag.
-
-Para tal é necessário adicionar uma etiqueta (tag) de 4 octetos , designada por etiqueta - Q (Q tag), a seguir ao campo de endereço fonte da trama 802.3.
-Vamos então analisar a trama utilizada neste standard, apresentada na imagem ~\ref{fig:TramaQ}.
+The BLE radio operates at the 2.4GHz band and employs a frequency hopping transceiver to combat interference and fading.
+It also employs two multiple access schemes: FDMA used to separate the 40 available physical channels, 37 of them are used as data channels and the remaining as advertising channels and TDMA in a polling scheme that is used when one device transmits a packet at a predetermined time and a corresponding device responds with a packet after a predetermined interval.
 
@@
 	\centering
-	\includegraphics[width=1\columnwidth]{./802_1Q}
-	\caption{Trama utilizada no IEEE802.1Q}
-	\label{fig:TramaQ}
+		\includegraphics[width=0.5\linewidth]{figures/BLEArchitecture.png}
+	\caption[Bluetooth Low Energy Architecture]{Bluetooth low energy architecture}
+	\label{fig:BLEarchitecture}
 \end{figure}
 
-\textbf{Identificador do protocolo VLAN:} Valor de 16 bits utilizado para identificar o protocolo VLAN. Para  IEEE802.1Q trama com Q-tag o valor é 0x8100.
+When looking at the BLE's architecture , which can be seen at \ref{fig:BLEarchitecture}, the Generic Access Profile (GAP) is one of the most relevant since it's responsible for working in conjunction with Generic Attribute (GATT) to define the base funcionality of BLE devices.
+The services that are made available by the GAP are: BLE device discovery, connection modes, security, authentication, association models and service discovery.
+In addiction to this, it also defines four different roles to describe a device, allowing for the controllers to be optimized in funtion of the device's desired roles: 
+Broadcaster, role optimized for transmitter-only applications; 
+Observer, role optimized for receiver-only applications and it's complementary to the broadcaster role;
+Peripheral, role optimized for devices that only want to suppot a single connection, allowing for a much less complex controller due to the fact that it only needs to support the slave role and not the master one;
+Central, role supports multiple connections and funtions as the initiator for all of them.
+These connection are all made with Peripheral devices and its controller must support the master role in a connection and allow for more complex funtions, in comparison to the remaining roles.
 
-\textbf{Prioridade (PCP, Priority control point):} Campo de 3 bits que definem o nível de prioridade da trama.
-Este valor é utilizado para dar prioridades diferentes a certo tipos de data (voice,video,etc.)
+Another important component is the Attribute Protocol (ATT) Protocol which is responsible for implementing the Peer-to-peer(P2P) protocol between an attribute server and client.
+This communication happens in a dedicated fixed  channel and a server can send through it responses, notifications and indications, while the client can send requests, commands and confirmations.
+The ATT allows the clients to read and write values of attributes on a peer device acting as a server.
 
-\textbf{Identificador de formato canónico (CFI) :} Campo de 1 bit utilizado para o propósito de compatibilidade entre ethernet (valor 0)  e token ring (valor 1). Este campo já não é mais utilizado.
+The last component is the Generic Attribute (GATT) Profile which is responsible for creating a framework for the ATT, in which it's represented the funcionalities of an ATT server.
+This profile describes the hierarchy of services, characteristics and attributes existent in the server and provides an interface for discovering, reading, writing and indicating service characteristics and profiles.
 
-\textbf{Identificador VLAN:} campo de 12 bits que especifica a VLAN à qual a trama pertence.
+Bluetooth low energy devices utilize profiles which define the required functionalities of the device.
+It also defines application behaviour and data formats and as such when two devices comply with all the requirements of a Bluetooth profile, application interoperability is achieved.
+Each Bluetooth profiles describes its requirements necessary for devices to create a connection, to find available services and connection information required for making application level connections.
 
-Este standard está limitado em termos de escalabilidade pelo parâmetro identificador VLAN, visto que é composto somente por 12 bits.
-Assim sendo só é possível ter 4094 VLAN diferentes pois os valores 0x000, que indica que um frame não tem VLAN ID, e 0xFFF, reservado para implementação, não podem ser utilizados.
-Devido a esta limitação este standard pode funcionar nas restrições de uma só organização, mas rapidamente se torna obsoleto quando os “Service Providers” têm como objetivo distribuir serviços ethernet para múltiplos end-users numa infrastructura de rede partilhada.
-
-Tendo analisado o standard IEEE802.1Q podemos então analisar o seu funcionamento em função da imagem ~\ref{fig:VLANex}.
-
- Tendo analisado o standard podemos então visualizar o seu funcionamento através da figura x.
-
- Estamos então perante um cenário onde há três VLANs diferentes, distribuídas aleatoriamente na rede.
-Para cada uma das VLANs há um exemplo de transmissão de informação, representado com as setas de cores, que demonstra como é feita a comunicação entre estações na mesma VLAN.
-
- Vamos então seguir o exemplo de uma mensagem da VLAN1 representada pela cor verde.
-Neste caso uma estação na VLAN1 mais à esquerda decide enviar uma mensagem para toda as restantes estações da sua VLAN e para tal envia uma mensagem broadcast.
-A mensagem para as estações dentro da sua “nuvem” estão já recebidas e falta só chegar a todas as restantes na rede, para tal, a mensagem segue até ao seu switch associado S1 e este irá reencaminhar a mensagem para o resto da rede.
-O standard IEEE802.1Q garantiu que será o switch S1 a realizar o Q-tagging, isto é, inserir a tag 1 (cor verde) na trama enviada, como pode ser visto na representação da trama na figura X.
-A trama é encaminhada para o switch principal que segundo a sua tabela de encaminhamento irá transmitir a trama para todos os switches que tenham a VLAN1 nos seus portos, ou seja os switches S3 e S4.
-Quando a mensagem chega aos switches S3 e S4, eles retiram a tag e reencaminham a mensagem para as estações da VLAN1 que eles conhecem.
-
- \begin{figure*}
+\begin{figure}
 	\centering
- 	\includegraphics[width=0.75\textwidth,height=6cm]{./ad_ex}
- 	\caption{Exemplo de rede para IEEE802.1ad}
- 	\label{fig:adex}
- \end{figure*}
-
-\section{Standard IEEE802.1ad}
-\label{sec:802.1ad}
-
-Devido às limitações nas capacidades de escalonamento criou-se o standard IEEE802.1ad, também conhecido como "QinQ".
-Para analisar as inovações apresentadas em comparação com o seu standard antecessor, IEEE802.1Q, basta olhar para a figura ~\ref{fig:trama_ad}.
-
-Como é possível visualizar na figura, a grande diferença entre os standards foi a adição de uma segunda “Q-TAG” (campo apresentado no capitulo ~\ref{sec:802.1Q} com o IEEE802.1Q) entre o customer tag e o endereço de origem e a distinção entre os dois agora existentes.
-O antigo campo de 4 bytes ficou conhecido como o Customer Tag e o novo campo como Service Tag.
-
-\begin{figure}[htp]
-	\centering
-	\includegraphics[width=1\columnwidth]{./trama_ad}
-	\caption{Trama utilizada no IEEE802.1ad}
-	\label{fig:trama_ad}
+		\includegraphics[width=0.5\linewidth]{figures/profile.png}
+	\caption[Gatt-based profile hierarchy]{Gatt-based profile hierarchy}
+	\label{fig:profile}
 \end{figure}
 
-Este ID, Service tag, é utilizado para identificar o serviço na rede fornecedora, enquanto que o VLAN ID do cliente, designado como C-tag, permanece intacto e não é alterado pelo fornecedor do serviço.
-Isto permite ao C-tag ser transparente na rede.
-Os Provider Bridges utilizam o S-tag para identificar a que serviço as frames de um cliente pertencem, logo, cada serviço requer um S-tag distinto.
-Este standard foi criado por algumas razões importantes:
+The base profile that any Bluetooth system needs to include is the GAP.
+From this point, any additional profile implemented will be a superset of GAP, where GATT is included and specifies the profile hierarchy, or the structure in which profile data is exchanged.
+Figure  ~\ref{fig:profile} shows the hierarchy in a Gatt-based profile, with the profile being the top level and services and characteristics below.
+A profile is composed by one or more services.
+A service is a collection of data and associated behaviors to accomplish a particular function or feature of a device or portions of a device.
+A service is composed of characteristics and/or references to other services.
+A Characteristic is a value that is used in a service that has properties and configuration information that descrive how the value should be accessed as well as information on how to display the value.
+A characteristic is defined by its declaration, its properties, its value and may also be defined by its descriptor, which describes the value or permit configuration of the server relative to the value.
 
-\textbf{Aumento do limite de VLANs}: Este standard apresenta uma frame com duas tags passando a ter uma limitação teórica de 4096x4096=16777216 LAN’s, permitindo um acomodamento maior da rede.
+When looking at what's possible to achieve using the BLE technology there is the example of Apple's creation iBeacon which was presented in 2013 with the porpuse of implementing proximity sensing systems.
+The device is capable of playing on the broadcaster role and as such its objective is to send nearby compatible receivers certain information.
+Some examples of application are to track customers or trigger location-based actions on devices such as push notifications or checking in on social media, with pratical cases such as the usage of iBeacons by McDonalds to offer special offers to their customers in their fast-food stores.
+An indoor location system utilizing this technology was presented by Jingjing Yang et al \cite{ibeacon}, where these devices were used to indicate a pacient of his whereabouts through the proximity sensing proprieties and this information was later transfered over to a server in order to give clients a variaty of different services, from pacient counting, to nearby department's information and offer indoor guidance to the nearest available bed.
 
-\textbf{Compatibilidade:} compatível com o seu antecessor (802.1q) e, embora seja limitado a duas tags, não existe um tecto máximo no standard que limite um frame a mais do que duas tags, permitindo um crescimento no protocolo.É mais fácil para os fabricantes de equipamento de redes modificar o equipamento já existente criando múltiplos cabeçalhos 802.1q do que modificar o equipamento para implementar alguma hipotética extensão do cabeçalho da VLAN ID que não seja do formato 802.1q.
+When utilizing BLE for indoor location the usual metric used to calculate distances is the RSSI.
+This metric withint the context of bluetooth brings to surface several issues such as the fact that RSSI as a metric is very accurate only when the target is within a meter of the beacon, since the value decreases as the inverse of the square of the distance to the beacon .
+As such when developing solutions for indoor location that require system with high accuracy capable of tracking moving objects, the usage of RSSI can't be utilized without further work.
+Faragher et al \cite{bleacc} tackled one of the techniques used to improve BLE system's accuracy, fingerprinting, by verifying the effects caused by the device deployment density within the required location.
+This experiment also puts into evidence one of the downsides of the bluetooth technology being that its scalability is low, besides requiring higher density in order to increase accuracy, due to their low range any need to increase coverage leads to increased costs.
 
-Contudo este standard não corrigiu todas as limitações existentes apresentadas no capitulo XX.
+Cricket
+Active Bat
+Radar
+active badge
 
-Continua a não haver separação entre as MAC addresses do fornecedor e do cliente, o que leva a que possa haver um problema de segurança, uma vez que a informação do seu endereçamento está visível fora do domínio de segurança da rede.
-Outra das limitações é que cada switch tem de saber todos os “MAC addresses”, o que implica que cada switch Provider Bridge deverá ter capacidade para suportar centenas de milhares de endereços, o que vai introduzir problemas de escalabilidade.
-Um exemplo representativo deste problema é no caso em que queremos adicionar um novo utilizador à rede, sendo que tem que ser memorizado pelos switches do fornecedor caso contrário, quando ocorrer uma falha na rede do consumidor, a ação tomada pelo Spanning Tree Protocol (STP) pode afectar a rede do consumidor.
-Os "downtimes" da rede em caso de falham continuam dependentes da execução completa da STP, mantendo assim as limitações que não permitem que a ethernet seja utilizada nas redes metropolitanas.
+recent stuff
 
-Podemos então passar a visualizar a figura ~\ref{fig:adex} de modo a perceber em termos funcionais o que foi alterado.
-Começamos por denotar que existem dois tipos diferentes de switches, os customer bridges (CB), switches que utilizam o IEEE802.1Q e os Provider bridges (PB), que funcionam com o IEEE802.1ad agora exemplificado.
+In 2011, LifeMap \cite{lifemap} introduced one of the first to attempt to track indoor location with unconstrained phone placement. Not Bluetooth
 
-Um edge switch, nome atribuído aos switches que estão na fronteira, isto é, comunicam com customer bridges, tem a funcionalidade de inserir e remover a service tag das tramas que por eles passam, ou seja, ao receber uma trama vinda de um CB é inserida uma S-tag correspondente ao serviço associado à trama.
-De seguida a trama será reencaminhada dentro desta nuvem que engloba os PBs todos, chegando a um switch VLAN-aware.
-Esta característica usada para descrever certos switches, indicam que o switch em causa conhece toda a rede e é capaz de reencaminhar correctamente a trama para o seu destino.
-Ao chegar de novo a um edge switch, o mesmo irá retirar a S-tag da trama e reencaminhar para o CB adjacente.
+Zonith \cite{zonith} introduced a bluetooth based location system with the objective of tracking the position of workers in dangerous environments.
+Any device registered in the zonith implemented network would be continuously tracked and accounted for in each of the system's functionalities such as, sounding an alarm whenever a lone worker doesn't move or responde within a time interval (Lone worker protection) or providing a quick an precise location of any worker that has requested for help.
+This system's installation requires planing of the best locations to place the beacons and number required of beacons in order to be able to provide enough courage and make sure the system provides the required quality.
 
-\section{Standard IEEE802.1ah}
-\label{sec:802.1ah}
+lifemap
+zonith
 
-O IEEE802.1ah, denominado de Provider Backbone Bridges (PBB) ou MAC-in-MAC, sucessor do standard apresentado no último capitulo evoluiu a ethernet ao adicionar endereços MAC dedicados para os service providers.
-Pelo figura ~\ref{fig:trama_ah} é possível visualisar todas as alterações que foram feitas com este standard:
+Chen et al. [2010] present an inquiry-based locating approach using Bluetooth
+RSS measurements. There also exist several commercial Bluetooth-based indoor positioning
+systems, such as ZONITH [Teldio 2006], which is used to assist people who
+work in hazardous environments.
 
-\begin{figure}[htp]
+- BLE Projects
+
+- Architecture similar stuff?
+
+\section{Implementation}
+\label{sec:struture}
+
+The solution presented in this paper was made with the objective of presenting a generic indoor location system using bluetooth low energy.
+The system's architecture is presented in figure ~\ref{fig:architecture} and is divided three parts: the device, the application and the server.
+For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
+
+\begin{figure}
 	\centering
-	\includegraphics[width=1\columnwidth]{./ah_trama}
-	\caption{Trama utilizada no IEEE802.1ah}
-	\label{fig:trama_ah}
+		\includegraphics[width=0.5\linewidth]{figures/profile.png}
+	\caption[Gatt-based profile hierarchy]{Gatt-based profile hierarchy}
+	\label{fig:architecture}
 \end{figure}
 
-\textbf{endereço de destino backbone}: campo de 6 bytes que contem o endereço MAC do switch PBB destino;
+\subsection{ BLE beacon}
+\label{subsec:beacon}
 
-\textbf{endereço de fonte backbone}: campo de 6 bytes que contem o endereço MAC do switch PBB fonte;
+\subsection{ Server}
+\label{subsec:beacon}
 
-\textbf{identificador de VLAN backbone( B-VID):} campo de 2 bytes utilizado para identificar o indentificador de VLAN atribuido à trama
+\subsection{ Application}
+\label{subsec:beacon}
 
-\textbf{identificador de serviço:} campo de 3 bytes utilizador para indicar o serviço da trama
+- Overview of Architecture
 
-Todas estas adições no formato da trama permitiram ao IEEE802.1ah:
+- Location Provider
+	- BLE Search
+	- GATT Comunication
+	- Server Calculation
+	- Voting System  and CoO
 
-Oferecer a completa separação entre os MAC addresses do fornecedor e do cliente;
-Não impõr qualquer alteração ao processo de switching da Ethernet nas pontes fundamentais;
-Proporcionar uma clara separação entre o domínio do consumidor e do fornecedor;
-Deduzir os MAC addresses do consumidor apenas através dos Backbone Edge Bridges;
-Suportar até 16 milhões de VLAN’s (campo service-id de 24 bits).
+- Google Maps
+	- Indoor maps 
+	- Map insertion possibility
 
-Contudo, este standard também tem desvantagens, e uma das mais evidentes é o facto das redes PBB continuarem a utilizar o Spanning Tree Protocol para eliminar malhas fechadas.
+Images:
+	- Architecture Overview
+	- Location Provider Diagram
+	- App Screen
 
-\section{Standard IEEE802.1ag}
-\label{sec:802.1ag}
+\section{Performance Analysis}
+\label{sec:performance}
 
-o standard IEEE802.1ag, uma actualização significativa do IEEE802.1ah, será o último a ser apresentado pois foi com ele que se obteve a chamada carrier ethernet.
-Baseado na trama utilizada pelo IEEE802.1ah e sem necessidade de adição de novos campos apresentou um número elevado de inovações que permitiu finalmente que a ethernet fosse utilizada como uma opção viável nas redes metropolitanas.
-
-\textbf{Engenharia de tráfego:} tomando vantagem do facto que ao desligar algumas das funcionalidades, o hardware existente ethernet é capaz de ter um novo comportamento de reencaminhamento, ou seja, é capaz de realizar um reencaminhamento de ligação orientada.
-
-Os ethernet switches baseiam-se nos 60 bits , endereço MAC (48 bits) mais o identificador VLAN (12bits), para reencaminhar mensagens, sendo que os dois parâmetros são únicos globalmente.
-Visto que  o identificador VLAN é utilizado para identificar um domínio sem loops no qual os endereços MAC são "flooded", se optarmos por configurar rotas sem loops com endereços MAC, o identificador VLAN é libertado das suas funções.
-Assim sendo no IEEE802.1ag este parâmetro é utilizado para identificar rotas específicas na rede para um dado endereço MAC destino.
-
-\begin{figure}[htp]
-	\centering
-	\includegraphics[width=1\columnwidth]{./te_ag}
-	\caption{Visualização da engenharia de tráfego}
-	\label{fig:te_ag}
-\end{figure}
-
-Este standard permite que certos endereços tenham as suas tabelas de encaminhamento preenchidas pelo plano de controlo em vez de ser pela utilização do STP.
-O funcionamento de cada switch mantem-se intacto, a diferença é que já não há necessidade de mecanismos de aprendizagem, resultando em rotas predeterminadas pela rede e um comportamento totalmente previsível da rede sobre qualquer circunstância.
-A figura ~\ref{fig:te_ag} permite visualizar a engenharia de tráfego aplicada à rede no sentido em que é visível duas rotas para o mesmo endereço MAC de destino, o working path que representa o caminho habitualmente utilizado e o Protection Path (indicado por um identificador VLAN diferente) que é utilizado em caso de falha da rota principal.
-Esta implementação permitiu reduzir drasticamente o tempo de baixa da rede para situações de falha de rota, visto que ao invés de ser necessário esperar pelo STP bastanta reencaminhar o tráfego pelo working path.
-
-\textbf{Operações, administração e gestão (OAM):} Funcionalidades OAM estão bem definidas nas redes TDM e são um factor importante para garantir que as operadoras pode oferecer serviços de qualidade "carrier".
-Os OAMs estão divididos em duas áreas, gestão de falha e monitorização de performance ,vamos então ver o que o IEEE802.1ag melhorou nestes campos:
-
-\textbf{Gestão de falha:}Garante que quando ocorre uma falha na rede, o problema é encaminhado para o operador que depois toma a decisão mais correcta. Está divida em três partes:
-
-\textbf{deteção de falha:} Mensagens de verificação contínua (continuous check messages (CCMs)) são enviadas pela fonte para o destino em intervalos fixos de tempos.
-Se uma resposta não for recebido dentro de um período específico de tempo (normalmente 3 vezes o intervalo de envio), anuncia-se uma falha na rede.
-O limite de tempo imposto para recuperação da rede nas redes metropolitanas é de 50ms, este valor era impossível de obter com o STP.
-Com as CCMs ao definir o intervalo de envio em valores inferior ao limite, é possível detectar uma falha em menos de 50ms.
-
-\textbf{verificação da falha:} Mensagens de loopback (LBM) e respostas loopback (LBR) permitem descobrir se uma determinada rota funciona.
-Normalmente utilizadas após a criação de uma rota ou quando se detecta um falha, para verificar que ela ocorre entre dois endpoints.
-
-\textbf{isolação de falha:} Mensagens de link trace (LTM) e respostas link trace (LTR) permitem descobrir qual a ligação de uma certa rota que falhou.
-Ao enviar esta mensagem (LBM) numa determinada rota que contem uma falha, todos os switches que a receberem devem responder com uma LBR, deste modo é possível identificar a partir de que switch a mensagem parou de ser encaminhada, identificando a ligação partida.
-
-\textbf{Monitorização de performance:} As tramas têm diferentes níveis de performance visto que num serviço podem sofrer de delays devido a congestionamento, ou mesmo perdas de tramas.
-Serviços de vídeo e de voz são os mais susceptíveis  aos efeitos de latência e jitter.
-Assim sendo as redes de carrier ethernet requerem monotorização de performance avançadas de modo a ajudar os provedores de serviço a cumprir os requerimentos da rede.
-
-\textbf{rácio de tramas perdidas:} Conhecido como Frame loss ratio, é calculado usando CCMs e contadores.
-Cada um dos endpoints envia um certo número de CCMs e conta quantas é que foram recebidas.
-No fim os contadores são trocados para calcular a percentagem de tramas perdidas.
-
-\textbf{atrasos nas tramas:} Mensagens são enviadas com os tempos de envio.
-O switch destino pode então calcular o tempo que a trama demorou até ser recebida.
-Só funciona se ambos os switches tiverem os relógios sincronizados.
-
-\textbf{ variação dos atrasos nas tramas:} Mesmo conceito que o apresentado anteriormente excepto que são armazenados os valores dos delays de modo a ver a sua variação.
-
-\section{Comparação de eficiência}
-\label{sec:eficiencia}
-
-Até este ponto foram analisados todos os standard criados para a obtenção da carrier ethernet e os benefícios que apresentaram sem nunca ter focado nas contra-partidas.
-Com cada um destes standards houve necessidade de inserir mais informação na trama, aumentando a carga que tem que ser transportada com cada mensagem.
-Iremos então analisar os efeitos de cada standard na eficiência.
-
-A eficiência de um standard depende da seguinte fórmula:
-
-\begin{equation}
-	\label{eq:problem}
-	 \text {Eficiencia} = \frac{\text{tamanho  da informação útil}}{ \text{tamanho da trama}} \\ 
-\end{equation}\\
-
-Relembramos que a informação tem um valor variável entre 40 e 1500 bytes , como visto na secção XXX e que o tamanho da trama varia de standard para standard.
-Como ponto de partida para todas estas inovações estava a ethernet e portanto iremos analisar a sua eficiência como base de comparação.
-
-A trama ethernet apresentada na figura ~\ref{fig:EtherFrame} tem necessidade de 38 bytes de overhead que são sempre adicionados à data.
-Assim sendo podemos visualizar a variação da sua eficiência na figure ~\ref{fig:ef_802.3}.
-
-\begin{figure}[htp]
-	\centering
-	\includegraphics[width=1\columnwidth]{./ef_802_3}
-	\caption{ eficiência do IEEE802.1Q}
-	\label{fig:ef_802.3}
-\end{figure}
-
-Para os standards analisados neste paper, as suas eficiências são comparadas na figura ~\ref{fig:ef_all}.
-O standard IEEE802.1Q está representado com a cor azul e é o mais eficiente em comparação aos restantes devido a só ter sido adicionado um campo de 4 bytes na trama, totalizando a um overhead de 42 bytes.
-De seguida está o IEEE802.1ad com a cor vermelha, standard que adicionou um segundo campo de 4 bytes, para um overhead de tamanho 46 bytes.
-Por último está o IEEE802.1ah com a cor verde, que apresenta a pior eficiência de todos os standards apresentados devido ao tamanho total do overhead de 67 bytes que impõe às suas tramas.
-
-\begin{figure}[htp]
-	\centering
-	\includegraphics[width=1\columnwidth]{./ef_all}
-	\caption{eficiência do IEEE802.1Q,IEEE802.1ad e IEEE802.1ah}
-	\label{fig:ef_all}
-\end{figure}
-
-\section{Conclusions}
-\label{sec:conclusion}
-
-A ethernet como foi estudada no capítulo ~\ref{sec:Ethernet} tinha vários defeitos que a impediam de ser implementada nas redes metropolitanas.
-Através de todas as inovações que foram implementas, chegando ao nível apresentado no IEEE802.1ag, já é possível.
-De momento a ethernet já não está limitada pelos seus problemas de escalabilidades, apresenta altos níveis de segurança e permite obter um rede resiliente capaz de monotorização de performance e gestão de falhas, cumprindo assim os requisitos em termos de OAMs
+\section{Future Work}
+\label{sec:future}
 
@@
 
-\bibitem{ether}
- \texttt{http://listserv.linguistlist.org/pipermail/ads-l/1999-July/001027.html}
-\bibitem{Carrier}
- \texttt{Ethernet as a Carrier Grade Technology: Developments and Innovations} Rafael Sánchez,2008
- \bibitem{slides}
- \texttt{Slides da unidade curricular- Redes de Telecomunicações- Capitulo 3} João Pires
+\bibitem{surveygps}
+ \texttt{Hakan Koyuncu, Shuang Hua Yang ;A Survey of Indoor Positioning and Object Locating Systems http://paper.ijcsns.org/07_book/201005/20100518.pdf}
+
+\bibitem{surveywireless}
+ \texttt{Hui Liu; Survey of Wireless Indoor Positioning Techniques and Systems http://www.pitt.edu/~dtipper/2011/Survey1.pdf}
+
+\bibitem{reviewtechniques}
+ \texttt{Zahid Farid, Rosdiadee Nordin, and Mahamod Ismail; Recent Advances in Wireless Indoor Localization Techniques and System}
+
+\bibitem{survey2}
+ \texttt{Hakan Koyuncu, Shuang Hua Yang; A Survey of Indoor Positioning and Object Locating Systems; http://paper.ijcsns.org/07_book/201005/20100518.pdf}
+
+\bibitem{survey1}
+ \texttt{Jiang Xiao, Zimu Zhou,Youwen Yi and Lionel M. Ni; A Survey on Wireless Indoor Localization from the Device Perspective; http://www.tik.ee.ethz.ch/~zzhou/paper/csur16-xiao.pdf}
+
+\bibitem{blecore}
+\texttt{ }
+
+\bibitem{ibeacon}
+\texttt{ Jingjing Yang, Zhihui Wang and Xiao Zhang ;An iBeacon-based Indoor Positioning Systems for Hospitals; http://www.sersc.org/journals/IJSH/vol9_no7_2015/16.pdf}
+
+\bibitem{bleacc}
+\texttt{ R. Faragher and R. Harle; An Analysis of the Accuracy of Bluetooth Low Energy for Indoor Positioning Applications; http://www.cl.cam.ac.uk/~rmf25/papers/BLE.pdf}
+
+\bibitem{lifemap}
+\texttt{Yohan Chon and Hojung Cha; LifeMap: A SmartphoneBased Context Provider for Location-Based Services; https://pdfs.semanticscholar.org/cb50/e7344494f8b8fdfeba160f814cbfe3ce3356.pdf}
+
+\bibitem{zonith}
+\texttt{http://media.teldio.com/pdfs/Teldio_Indoor%20Positioning%20White%20Paper_v2.1a.pdf}
 
@@
 \end{document}
-This is never printed

### 2016-11-11 paper.tex
section 3 started
missing google maps integration and finishing mock provider

@@
 \section{Introduction}
-\label{sec:intro}
+\label{sec:Introduction}
 
@@
 
-Cricket
-Active Bat
-Radar
-active badge
+-Cricket
+
+-Active Bat
+
+-Radar
+
+-active badge
 
@@
 
-lifemap
-zonith
-
-Chen et al. [2010] present an inquiry-based locating approach using Bluetooth
-RSS measurements. There also exist several commercial Bluetooth-based indoor positioning
-systems, such as ZONITH [Teldio 2006], which is used to assist people who
-work in hazardous environments.
-
 - BLE Projects
@@
 The solution presented in this paper was made with the objective of presenting a generic indoor location system using bluetooth low energy.
-The system's architecture is presented in figure ~\ref{fig:architecture} and is divided three parts: the device, the application and the server.
+The system's architecture is presented in figure ~\ref{fig:architecture} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:app} alongside figures that show the functional prototype.
 For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
@@
 	\centering
-		\includegraphics[width=0.5\linewidth]{figures/profile.png}
-	\caption[Gatt-based profile hierarchy]{Gatt-based profile hierarchy}
+		\includegraphics[width=1\linewidth]{figures/Architecture.png}
+	\caption[System's Architecture]{System's Architecture}
 	\label{fig:architecture}
@@
 
+The beacons that were utilized are Texas Instruments CC2650STK devices which can be visualized in figure ~\ref{fig:beacon}.
+Alongside the device, which comes with a pre-installed bluetooth low energy program capable of giving information on each of its ten sensors through its predefined profiles, there is a texas smartphone application that can connect to a single device and read from its sensors.
+By using the texas Code Composer Studio (CSS), the pre-defined ble profile existant on the device could be altered.
+Upon further analysis of the profile, a characteristic was found for which the Universal Unique Identifier (UUID) of the service and the characteristic itself was found and as such this was the one that ended up being used to store the device's owner server's address.
+Since the device was already set to work as a pheripheral and it now stored the information relevant to the system, there was no need to do further work.
+
+\begin{figure}
+	\centering
+		\includegraphics[width=0.5\linewidth]{figures/beacon.jpg}
+	\caption[TI cc2650stk sensortag]{TI cc2650stk sensortag}
+	\label{fig:beacon}
+\end{figure}
+
 \subsection{ Server}
-\label{subsec:beacon}
+\label{subsec:server}
+
+The webserver was implement in Python 3 programming language.
+The program implements a simple tcp server capable of receiving multiple request at the same time.
+Each request starts with information sent from an application which include a pair of MAC address and associated RSSI value for each ble device that the same application found.
+Afterwards the list of pairs is filtered in order to remove any existant devices that are not present in the server's database of devices.
+
+Each server has a database that includes only ble devices.
+An entry (description of a device) in this database is composed by the device's mac address, its longitude and latitude and its building, floor and room name.
+In addition to the database, a server when initiated can store additional location info such as the server's street, number, zipcode, city and country, allowing this information to be transmited to the client in order to offer an additional level of location description to the user.
+The whole location specific information can be visualised in figure ~\ref{fig:AppMenu}.
+
+Upon having filtered the initial list of pairs, the Cell of Origin (CoO) technique is applied by verifying which of the devices produced a stronger signal on the receptor.
+Upon obtaining the closest device an answer is sent to the application containing all of the information associated to the server and the selected device.
+
+- Describe the Database ( insert image example of database with a few entries?? )
+
+- Mention capability to Insert aditional info, at the moment it displays geocoding on pop up menu in figure ~\ref{fig:AppMenu}.
 
 \subsection{ Application}
-\label{subsec:beacon}
+\label{subsec:app}
 
-- Overview of Architecture
+The Smartphone application was developed for Android using the Android Studio IDE.
+The Application is divided in two primary functional blocks, the Mock location Provider and the Google Maps Integrated Display, as can be visualized in figure ~\ref{fig:Architecture}.
 
-- Location Provider
-	- BLE Search
-	- GATT Comunication
-	- Server Calculation
-	- Voting System  and CoO
+The Mock Location Provider is implemenented if it was a Location provider, such as gps.
+The application functions works as a listener to a Location provider, in this case it listens to the Mock Provider that was implemented.
+By implementing the whole process of obtaining a location inside a service (the mock provider) , a new level of abstraction is added to the application.
+As such, whenever the application is signaled to obtain the user's current location, a request is made to the associated location provider and the application only need to listen for the answer that eventualy arrives.
 
-- Google Maps
-	- Indoor maps 
-	- Map insertion possibility
+The Mock Location Provider incorporates the first three steps present in figure ~\ref{fig:MockProvider}, which will now be explored individualy.
+The first step indicates the gathering of information of the surroundings of the user's device.
+When a request is made to the provider, a scan for nearby bluetooth low energy devices is made which will put the smartphone in a state of listening for incoming ble advertisement packets for half a second ( NEED TO VERIFY VALUE AND JUSTIFY, THERE WAS A PAPER ON THIS).
+During this scanning period, each time a device is found, the advertisement is registered in a list, which have a duplication prevention mechanism implemented.
+Once the period is over, the provider has avaliable a list of all the ble devices within range.
 
-Images:
-	- Architecture Overview
-	- Location Provider Diagram
-	- App Screen
+The second step involves taking the created list of devices, obtain a server adress and forward the same list to it.
+Once the first step is completed, the provider will analyze each entry at a time.
+For each device the provider will attempt to respond to the caught advertisement packet, resulting in a created connection.
+Once the connection is created the provider asks for the available services of the paired device.
+Upon receiving an answer, the list of services is swoop while looking for the service with the wanted UUID.
+If the device doesn't have the UUID that the provider is looking for, it can assume that the paired ble is not a beacon of our system, as such the connection is terminated.
+When the provider identifies that the device has the system's UUID, it requests the device to provide the service's existant characteristics.
+The provider will receive a list composed of the service's characteristics and it will search in it for the system's characteristics UUID, the one which contains the device's server's address.
+This search has the objective of confirming that the service existant in this device is indeed the one that was implemented for the system and not a device with another service that happened to have the same UUID.
+For any service outside those that are documented in the Bluetooth Special Interest Group (SIG), who have a specific UUID attached to them, the UUID is generated randomly and as such there is a small chance of collision.
+Once the wanted characteristic is found, the provider requests the device to read its value and stores the received value in a list.
+This list will contain the servers of the devices that were found, and for each address there will be a list corresponding to each device , and their corresponding rssi values, from the same owner.
+In order to quicken the previously described process, the provider keeps in cache the most recent contacted devices.
+Before attempting a connection, the provider confirms that the device isn't found in cache and when finishing a process, the associated device is inserted into the cache.
+
+When every device has been contacted, a voting system is actioned which will decide from the list of servers which one it will send the collected information to.
+The voting system uses an exponential function in order to attribute a weight to each server.
+INSERT FUNCTION AND EXPLAINATION.
+
+The voting system was implemented with the objective providing a thin security layer by allowing multiple devices of the same server to overcome a single attacker's device which happened to be close to the user.
+After obtaining each server's values, the one with the highest value is chosen and sent the list with all the devices.
+
+The Third step... CONTINUE
+
+-Explain step two: Ask each of found devices for their Owner; Explain the whole Communication process ( create image with the message process??
+) The process would be, connect -> request services -> Find wanted service -> Request Characteristics -> Find Characteristic wanted -> Read Value; Mention Cache properties available to speed up this step; Server Calculation of which to Contact ( the 2^X formula used to obtain most valuable, Voting system).
+
+- Explain step three:  TCP communcation; Formulate The message and send; Obtain Answer -> Gather Aditional Info (that which is used in PopUp menu) -> Return a Location to the APP
+
+One of the current limitations of the android API that deals with BLE is that the interface on the smartphone that is used to connect with a device has a fixed timeout time.
+MIGHT JUST NEED TO BE IMPROVED
+
+\begin{figure}
+	\centering
+		\includegraphics[width=0.5\linewidth]{figures/RequestLocation.png}
+	\caption[Mock Location Provider Workflow]{Mock Location Provider Workflow}
+	\label{fig:MockProvider}
+\end{figure}
+
+- Google Map Integration:
+
+- Explain Set four: We use the Mock Location Provider obtained Location to focalise the camera, insert location marker and update PopUp Menu
+
+- Possible: Study Google Maps Possibility of load our own maps ( i believe it's possible, need to check api), this would explain scalability Map wise.
+
+- The Final result is the following two images, figure ~\ref{fig:AppFocus} and figure ~\ref{fig:AppMenu}.
+
+\begin{figure}
+	\centering
+		\includegraphics[width=0.5\linewidth]{figures/app_focused.png}
+	\caption[Application screen showing a focused location on a room]{Application screen showing a focused location on a room}
+	\label{fig:AppFocus}
+\end{figure}
+
+\begin{figure}
+	\centering
+		\includegraphics[width=0.5\linewidth]{figures/app_focused_menu.png}
+	\caption[Application screen showing additional information of location]{Application screen showing additional information of location}
+	\label{fig:AppMenu}
+\end{figure}
 
@@
 
+- Energy Consumption in general
+
+- Energy consumption vs Cycle time vs Accuracy?
+
+- Location Accuracy
+
+- Time for request completion
+
+- Data Requiremnts?
+
 \section{Future Work}
 \label{sec:future}
+
+- Possible Improvement: Paralel ble connections; Improvements on lateration algorithm;  Rest depends on the Performance results
+
+- Conclusions
 
@@
 \bibitem{zonith}
-\texttt{http://media.teldio.com/pdfs/Teldio_Indoor%20Positioning%20White%20Paper_v2.1a.pdf}
+\texttt{http://media.teldio.com/pdfs/Teldio_Indoor}
 

### 2016-11-14 paper.tex
finished implementation

@@
 
+TODO Existante work
+
 -Cricket
@@
 
-The Mock Location Provider is implemenented if it was a Location provider, such as gps.
+The Mock Location Provider is implemenented as if it was a Location provider, such as gps.
 The application functions works as a listener to a Location provider, in this case it listens to the Mock Provider that was implemented.
@@
 
+\begin{figure}
+	\centering
+		\includegraphics[width=0.5\linewidth]{figures/RequestLocation.png}
+	\caption[Mock Location Provider Workflow]{Mock Location Provider Workflow}
+	\label{fig:MockProvider}
+\end{figure}
+
 The Mock Location Provider incorporates the first three steps present in figure ~\ref{fig:MockProvider}, which will now be explored individualy.
 The first step indicates the gathering of information of the surroundings of the user's device.
-When a request is made to the provider, a scan for nearby bluetooth low energy devices is made which will put the smartphone in a state of listening for incoming ble advertisement packets for half a second ( NEED TO VERIFY VALUE AND JUSTIFY, THERE WAS A PAPER ON THIS).
+When a request is made to the provider, a scan for nearby bluetooth low energy devices is made which will put the smartphone in a state of listening for incoming ble advertisement packets for half a second ( NEED TO VERIFY VALUE AND JUSTIFY, THERE WAS A PAPER ON THIS TODO).
 During this scanning period, each time a device is found, the advertisement is registered in a list, which have a duplication prevention mechanism implemented.
@@
 
-The Third step... CONTINUE
-
--Explain step two: Ask each of found devices for their Owner; Explain the whole Communication process ( create image with the message process??
-) The process would be, connect -> request services -> Find wanted service -> Request Characteristics -> Find Characteristic wanted -> Read Value; Mention Cache properties available to speed up this step; Server Calculation of which to Contact ( the 2^X formula used to obtain most valuable, Voting system).
-
-- Explain step three:  TCP communcation; Formulate The message and send; Obtain Answer -> Gather Aditional Info (that which is used in PopUp menu) -> Return a Location to the APP
+The Third step involves a simple client/server tcp interaction.
+The application starts off by formulating the message that it will later on send to the server, this message includes all pairs of device mac address and its associated rssi value captured by the application on the first step.
+Once the message is computed, the application attemps to create a connection with the server at the chosen address at the end of step two.
+With the connection established, the message is forwarded to the server and the application is put onto blocked state where it awaits for an answer.
+Upon arrival, the answer received is checked for valid location, its information is process and the connection is terminated.
+The information contained inside the received message, which was described in section ~\ref{subsec:server}, is then processed into the adequate class capable of storing a geographic location and the same is broadcasted from the mock location provider to its listener.
 
 One of the current limitations of the android API that deals with BLE is that the interface on the smartphone that is used to connect with a device has a fixed timeout time.
-MIGHT JUST NEED TO BE IMPROVED
+MIGHT JUST NEED TO BE IMPROVED TODO
 
-\begin{figure}
-	\centering
-		\includegraphics[width=0.5\linewidth]{figures/RequestLocation.png}
-	\caption[Mock Location Provider Workflow]{Mock Location Provider Workflow}
-	\label{fig:MockProvider}
-\end{figure}
+The Google Maps Integrated display is implemented using the Google Maps Android API.
+By using Google Maps it was possible to aliviate the weight on application since there wasn't need implement file transfer of indoor building's maps from each dedicated server to each request, which aliviated the servers aswell since there was no need to store its associated building's maps on it.
+Managing the maps was something that was aswell fortunately unnecessary and as such all these features were provided by google maps service.
+By making this development choice, the system as whole became closer to the desired generic approach while making possible for seamless transition between indoor/outdoor maps.
+The only imposed restriction is related to the addiction of new indoor maps onto the google maps, which is possible and well documented but dependent on a third party.
 
-- Google Map Integration:
+The Fourth step is called when the application receives a proper location from the request made onto the location provider.
+With the device's location known, a marker is placed on the map with the obtain coordinates (longitude, latitude), the camera is moved in order to be centralized on the position and fully displaying the indoor level map, and the menu visible on figure ~\ref{AppMenu} is updated with the information that is bundled with the received location.
+In order to show the correct level on a multi level building, the "floor" information present in the menu is utilized.
+The API allows for obtaining a list of existant levels on which the maps' camera is focused and as such it's possible to find out to which level the provided locaiton belongs and make so that the application shows it.
 
-- Explain Set four: We use the Mock Location Provider obtained Location to focalise the camera, insert location marker and update PopUp Menu
+The pop-up menu was implemented to demonstrate the capacity of providing additional information associated with each location, be it geo-location taxonomy as it is currently implemented or possibly a description of the located room, an hyperlink of some sort or any other type of data that someone implemented this system would like to provide to its users.
 
-- Possible: Study Google Maps Possibility of load our own maps ( i believe it's possible, need to check api), this would explain scalability Map wise.
-
-- The Final result is the following two images, figure ~\ref{fig:AppFocus} and figure ~\ref{fig:AppMenu}.
+The final state of the implemented system can be visualized in figure ~\ref{fig:AppFocus} and figure ~\ref{fig:AppMenu}.
+The first displays the case of obtaining a location, where the marker has been placed and the camera zoomed
 

### 2016-12-17 paper.tex
updated
generic and such

@@
 
+Generic and extensible indoor location services
+
+Multimodal indoor location service/framework
+
 \title{{ \normalsize Master Thesis} \\
@@
 Indoor positioning systems have greatly evolved in the past few years due to the great success of its counter-part, the global positioning system (GPS), in outdoor environments but failure to reproduce the same results in indoor environments.
-Since GPS is an outdoor position systems and is based on a network of satelite, when the required scenario for position tracking is inside a building, new constraints are presented onto the process such as the attenuation and reflection of eletromagnetic waves upon collision with building walls and obstacles \cite{surveygps}.
+Since GPS is an outdoor position systems and is based on a network of satelite, when the required scenario for position tracking is inside a building, new constraints are presented onto the process such as the attenuation and reflection of eletromagnetic waves upon collision with building walls and obstacles \cite{survey1}.
 As such there was a need to find reliable indoor systems that by nature would already be able to heavily reduce the impact of some of the mentioned constraints.
@@
 In order to understand indoor position there is a need to understand the full scope of variables that come to surface when moving from outdoor to indoor.
-When developing a system there is a need to make sure that it can tackle the chalenges such as small space dimension which reforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques}.
+When developing an indoor system there is a need to make sure that it can tackle the chalenges such as small space dimension which reforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques}.
 All of the previously mentioned affect the way electromagnetic waves propagate in an indoor environment leading to problems related to severe multipath and reflection on existent surfaces \cite{surveywireless}.
@@
 
-The solution in this paper is based on the Bluetooth Low energy technology using the CoO method to obtain the device's location and as such for what technologies and location methods are concerned, only the used ones will be focused.
+When taking into consideration every possible technology and existant algorithm that are capable of being applied on indoor location, we find ourselves with a wide variaty of solutions.
+This high number of variations has led to an increase on the number of existing surveys which attemp to gather make a collection of what has been achieved in the field.
+Looking at this opportunity I saw the possibility of attempting to create a generic platform capable of being utilized by any existing indoor location system.
+In order to implement and test the solution there was a need to decide on which technology to be used and as such I opted by the bluetooth low energy.
 
-In this paper is structured in a way that in \ref{sec:related} an overview of Bluetooth low energy (BLE) and existent projects and their associated architecture are presented while shedding light on the benefits and limitations of the chosen technologies, in \ref{sec:structure} the proposed system is described based on technology and method used as well as the architecture of the would system, \ref{sec:performance} analyzes the different aspects of the presented solution in terms of energetic efficiency, accuracy and response timings, finalizing in \ref{sec:future} by overviewing what could be carried out in order to further develop the existing work and by concluding the paper.
-
-- Falta de framework, não tenho bases, vale a pena mencionar na introdução?
+In this paper is structured in a way that in \ref{sec:related} an overview of Bluetooth low energy (BLE) and existent projects and their associated architecture are presented while shedding light on the benefits and limitations of the chosen technologies.
+Section \ref{sec:architecture} overviews the architecture of the proposed generic system, while section \ref{sec:structure} presents the finalized system by analyzing the utilized technologies.
+Section \ref{sec:performance} analyzes the different aspects of the presented solution in terms of energetic efficiency, accuracy and response timings, finalizing in section \ref{sec:future} by overviewing what could be carried out in order to further develop the existing work and by concluding the paper.
 
@@
 \label{sec:related}
+
+Indoor location systems is a field that has only been researched for a few couple of years but from the very begining until now there has been a very large amount of changes.
+In order to understand the progresses made there is a need to dwell into the systems which functioned as bases for progress.
+
+In 1992 the Active system \cite{badge} was presented as an infrared solution which was capable of provided room-based position tracking, due to  infrared signals being unable to to travel through walls, and whose's information was then stored in a centralized location database \cite{badge1}.
+This system imposed users to carry with them an ID card equiped with an IR LED which emitted a signal each fifteen seconds.
+The long signal frequency was due to the chance of collision of signals from multiple users in the same location while also having in mind energy cost which need to be kept as low as possible.
+One of the downsides was the decrease in accuracy as the user's location is only known at best to a 15-second window.
+The user's location was obtained  through the implementation of a network of sensors which act as receivers and then forward the obtained information to the master station which would take care of polling the sensors, processing data and then making it available to users.
+One of the biggest issues that were raised with this system was the privacy concerns raised by the clients due to its object tracking nature.
+
+At the beggining of the twentieth century, two remarkable projects were presented, Bat system \cite{bat} and RADAR\cite{radar}.
+The first one, Bat Ultrasonic Location System, was a system capable of tracking various objects, each with a small wireless device called bat \cite{bat1, bat2}.
+A bat consisted of a radio transmitter, each had a unique identifier and was ment to be carried by personnel or attached to objects.
+In order to locate a bat, there was a base station which periodically transmitted a signal with a single identifier which caused the corresponding bat to emmit an ultrasound pulse.
+The same pulse was then captured by an existing grid of receivers, which were placed on the ceiling of the areas to be covered, which recorded the time of arrival of the ultrasound.
+Since the receivers are also capable of receiving the initial base station's message they are capable of computing the ultrasound's time of flight, which is then forwarded to the base station for further computation.
+Since the speed of sound in the air is known, the system was capable of converting the time of flight of the ultrasound and convert it into bat-receiver distance.
+In order to obtain the bat's location, if at least three receivers had caught the bat's signal, it was possible to utilize multilateration to obtain its 3D position.
+Due to the nature of the system, only one bat can be localised at a time, as such the process was diveded into time slots which garanteed that each time slot had enough time to obtain a bat's position and for its emitted signals to die out.
+
+The Radar system is a RF-based system to locate and track users inside buildings.
+The system is devided into two phases, an offline phase used for data collection and a real-time phase where a location is obtained.
+For data collection the mobile application and the base stations are synchronized, in order to be able to obtain time stamps and then the mobile user periodically emits UDP packets to the base stations (BS) .
+Each BS records the RF signal strength measurement together with the time stamp.
+This information gathering happens in both offline and real-time phases.
+During only the offline phase, while the mobile user emits packets, it also indicates his position on the map through (x,y) coordinates together with the timestamp.
+The real-time phase utilises the real-time user's packet associated RF signal strength, which is obtained at each bs and later on forwarded to the central computer where the computation is made, and examinates it in order to find the best fit for the current transmitter position.
+
+The last remarkable project, Cricket, was introduced in 2005 and managed to tackled some of the problems presented in the last mentioned systems \cite{cricket1}.
+Each node in the cricket system is composed of a RF transceiver and hardware capable of generating and receiving ultrasonic signals.
+There are two types of nodes, beacons, which are fixed reference points and are attached to the ceiling or walls of the building, and listeners which are attached to the objects that need to be tracked.
+Each beacon periodically transmits a RF signal message containing beacon specific information, such as unique ID and beacon position.
+Whenever a RF signal is transmitted, an ultrasonic pulse is also emitted which enables listeners to measure their distance to the beacons by using the time difference of arrival times of the RF and ultrasonic signals.
+Each listener utilizes the RF signal's beacon information alongside the obtain distances to beacons to compute their space position and orientation.
+Crickets Architecture allowed to solve the user pribacy, decentralized and ease of deployment issues there were present in the remaining projects \cite{cricket}.
+
+TODO recent stuff
 
@@
 \end{figure}
+
+When looking at what's possible to achieve using the BLE technology there is the example of Apple's creation iBeacon \cite{ibeacon} which was presented in 2013 with the porpuse of implementing proximity sensing systems.
+The device is capable of playing on the broadcaster role and as such its objective is to send nearby compatible receivers certain information.
+Some examples of application are to track customers or trigger location-based actions on devices such as push notifications or checking in on social media, with pratical cases such as the usage of iBeacons by McDonalds to offer special offers to their customers in their fast-food stores.
+An indoor location system utilizing this technology was presented by Jingjing Yang et al \cite{ibeacon1}, where these devices were used to indicate a pacient of his whereabouts through the proximity sensing proprieties and this information was later transfered over to a server in order to give clients a variaty of different services, from pacient counting, to nearby department's information and offer indoor guidance to the nearest available bed.
+
+When utilizing BLE for indoor location the usual metric used to calculate distances is the RSSI.
+This metric withint the context of bluetooth brings to surface several issues such as the fact that RSSI as a metric is very accurate only when the target is within a meter of the beacon, since the value decreases as the inverse of the square of the distance to the beacon .
+As such when developing solutions for indoor location that require system with high accuracy capable of tracking moving objects, the usage of RSSI can't be utilized without further work.
+Faragher et al \cite{bleacc} tackled one of the techniques used to improve BLE system's accuracy, fingerprinting, by verifying the effects caused by the device deployment density within the required location.
+This experiment also puts into evidence one of the downsides of the bluetooth technology being that its scalability is low, besides requiring higher density in order to increase accuracy, due to their low range any need to increase coverage leads to increased costs.
+
+Zonith \cite{zonith} introduced a bluetooth based location system with the objective of tracking the position of workers in dangerous environments.
+Any device registered in the zonith implemented network would be continuously tracked and accounted for in each of the system's functionalities such as, sounding an alarm whenever a lone worker doesn't move or responde within a time interval (Lone worker protection) or providing a quick an precise location of any worker that has requested for help.
+This system's installation requires planing of the best locations to place the beacons and number required of beacons in order to be able to provide enough courage and make sure the system provides the required quality.
+
+\section{Architecture}
+\label{sec:architecture}
+
+The solution presented in this paper was made with the objective of creating a generic indoor location system capable of being implemented using any existant indoor system created.
+The generic system's architecture is presented in figure ~\ref{fig:generic} and it's divided in 4 parts: beacon, location server, map server and smartphone application.
+
+\begin{figure}
+	\centering
+		\includegraphics[width=1\linewidth]{figures/generic.png}
+	\caption[Generic System's Architecture]{Generic System's Architecture}
+	\label{fig:generic}
+\end{figure}
+
+The beacon represents any form of hardware responsible for providing fixed reference points which are fundamental in calculating a user's position.
+A beacon needs to be be composed of two structural blocks: The Unique ID, which stores any information crucial to identifying a beacon, and the Hardware specific software component, which contains all the necessary software to accomplish communication with a certain technology.
+
+The location server is divided in three: the Communication block, which is responsible for creating, managing and terminating a connection with the mobile application; the location algorithm, architectural block responsible for applying a certain location algorithm to the data received from the application in conjuntion with the server's device information in order to obtain a concrete location; and the beacon manager which functions as a database where all the beacon associated with the server are stored.
+
+The map server represents the architectural block responsible for providing the maps associated to the location of user obtained through the location service on the application.
+As such it requires a communication block which needs to be capable of communication with the map service of the application and takes care of creating, managing and terminating connections, and a map manager which stores the existing maps that are to be transfered to the application.
+
+The smartphone application is divided into three functional blocks: core application, location service and map service.
+The core application contains the core functionality of the application and is in charge of communicating with the existing services.
+Communication with the location service allows for requesting the user's location which is then forwarded to the map service, in order to obtain the associated map.
+The location service is composed by four blocks: the API component through which it communicates with the core of the application; the tag reader which is the functional block responsible for reading a certain type of tags (beacons) and as such its structure will depend on the type of beacon utilized by the service; the tag decoder, functional block responsible of utilizing the information obtained through the tag reader and process it in order to be forwarded to communication block; the communication block which takes care of the creating, managing and terminating connections with a location server.
+
+It might be relevant to analyze some types of indoor location system that could diverge from this architecture, such as camera or Quick Response (QR) code-based systems.
+These systems would still follow the proposed architecture except for the beacon component which would not be existant and the location service's tag reader block wouldn't need any sort of communication since it would be responsible of utilizing the camera to capture either a fotograph or reading a QR core.
+
+\begin{figure}
+	\centering
+		\includegraphics[width=0.75\linewidth]{figures/altgeneric.png}
+	\caption[Adjustments to the application required for multiple indoor location systems]{Adjustments to the application required for multiple indoor location systems}
+	\label{fig:altgeneric}
+\end{figure}
+
+Although figure \ref{fig:generic} is representative of a system utilizing a single type of technology, the presented solution is scalable, allowing for insertion of additional indoor location systems onto a single architecture.
+The changes required for implementing such a scenario would be to add one extra layer of complexity on the android application as present on figure \ref{fig:altgeneric}.
+In this new architecture, the application's core would only communicate with the location manager and the latter would now be responsible of managing through all of the location services and request execution of whichever service would be more suited or available.
+The location manager would also be required to manage through the existant map services, which may or may not be in same number as the existant location services, and make sure that the data obtained from a certain location service is forwarded to the correct map service which must capable of processing it.
+
+\section{Implementation}
+\label{sec:struture}
+
+\subsection{Bluetooth Low energy}
+\label{ble}
 
@@
 
-When looking at what's possible to achieve using the BLE technology there is the example of Apple's creation iBeacon which was presented in 2013 with the porpuse of implementing proximity sensing systems.
-The device is capable of playing on the broadcaster role and as such its objective is to send nearby compatible receivers certain information.
-Some examples of application are to track customers or trigger location-based actions on devices such as push notifications or checking in on social media, with pratical cases such as the usage of iBeacons by McDonalds to offer special offers to their customers in their fast-food stores.
-An indoor location system utilizing this technology was presented by Jingjing Yang et al \cite{ibeacon}, where these devices were used to indicate a pacient of his whereabouts through the proximity sensing proprieties and this information was later transfered over to a server in order to give clients a variaty of different services, from pacient counting, to nearby department's information and offer indoor guidance to the nearest available bed.
-
-When utilizing BLE for indoor location the usual metric used to calculate distances is the RSSI.
-This metric withint the context of bluetooth brings to surface several issues such as the fact that RSSI as a metric is very accurate only when the target is within a meter of the beacon, since the value decreases as the inverse of the square of the distance to the beacon .
-As such when developing solutions for indoor location that require system with high accuracy capable of tracking moving objects, the usage of RSSI can't be utilized without further work.
-Faragher et al \cite{bleacc} tackled one of the techniques used to improve BLE system's accuracy, fingerprinting, by verifying the effects caused by the device deployment density within the required location.
-This experiment also puts into evidence one of the downsides of the bluetooth technology being that its scalability is low, besides requiring higher density in order to increase accuracy, due to their low range any need to increase coverage leads to increased costs.
-
-TODO Existante work
-
--Cricket
-
--Active Bat
-
--Radar
-
--active badge
-
-recent stuff
-
-In 2011, LifeMap \cite{lifemap} introduced one of the first to attempt to track indoor location with unconstrained phone placement. Not Bluetooth
-
-Zonith \cite{zonith} introduced a bluetooth based location system with the objective of tracking the position of workers in dangerous environments.
-Any device registered in the zonith implemented network would be continuously tracked and accounted for in each of the system's functionalities such as, sounding an alarm whenever a lone worker doesn't move or responde within a time interval (Lone worker protection) or providing a quick an precise location of any worker that has requested for help.
-This system's installation requires planing of the best locations to place the beacons and number required of beacons in order to be able to provide enough courage and make sure the system provides the required quality.
-
-- BLE Projects
-
-- Architecture similar stuff?
-
-\section{Implementation}
-\label{sec:struture}
-
-The solution presented in this paper was made with the objective of presenting a generic indoor location system using bluetooth low energy.
-The system's architecture is presented in figure ~\ref{fig:architecture} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:app} alongside figures that show the functional prototype.
+The implementation presented in this paper was created by utilizing the generic indoor location system presented in section \ref{sec:architecture} and applying it with bluetooth low energy.
+The system's architecture is presented in figure ~\ref{fig:implementation} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:app} alongside figures that show the functional prototype.
 For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
@@
 	\centering
-		\includegraphics[width=1\linewidth]{figures/Architecture.png}
+		\includegraphics[width=1\linewidth]{figures/System.png}
 	\caption[System's Architecture]{System's Architecture}
-	\label{fig:architecture}
+	\label{fig:implementation}
 \end{figure}
@@
 The Smartphone application was developed for Android using the Android Studio IDE.
-The Application is divided in two primary functional blocks, the Mock location Provider and the Google Maps Integrated Display, as can be visualized in figure ~\ref{fig:Architecture}.
+The Application is divided in two primary functional blocks, the Mock location Provider and the Google Maps Integrated Display, as can be visualized in figure ~\ref{fig:implementation}.
 
@@
 \begin{thebibliography}{9}
-
-\bibitem{surveygps}
- \texttt{Hakan Koyuncu, Shuang Hua Yang ;A Survey of Indoor Positioning and Object Locating Systems http://paper.ijcsns.org/07_book/201005/20100518.pdf}
-
 \bibitem{surveywireless}
@@
 
+\bibitem{survey3}
+\texttt{Klaithem Nuaimi and Hesham Kamel; A survey of indoor positioning systems and algorithms; https://www.researchgate.net/publication/224242434_A_survey_of_indoor_positioning_systems_and_algorithms}
+
 \bibitem{blecore}
@@
 \bibitem{ibeacon}
+\texttt{ibeacon insider: what's ibeacon; http://www.ibeacon.com/what-is-ibeacon-a-guide-to-beacons/}
+
+\bibitem{ibeacon1}
 \texttt{ Jingjing Yang, Zhihui Wang and Xiao Zhang ;An iBeacon-based Indoor Positioning Systems for Hospitals; http://www.sersc.org/journals/IJSH/vol9_no7_2015/16.pdf}
@@
 
+\bibitem{badge}
+\texttt{The Active Badge System; http://www.cl.cam.ac.uk/research/dtg/attarchive/ab.html}
+
+\bibitem{badge1}
+\texttt{WANT, R., HOPPER, A., FALCAO, V., AND GIBBONS, J.;The Active Badge Location System;  http://www.cl.cam.ac.uk/research/dtg/www/publications/public/files/tr.92.1.pdf}
+
+\bibitem{bat}
+\texttt{ The Bat Ultrasonic Location System ;http://www.cl.cam.ac.uk/research/dtg/attarchive/bat/ }
+
+\bibitem{bat1}
+\texttt{Andy Harter, Andy Hopper, Pete Steggles, Andy Ward and Paul Webster;The Anatomy of a Context-Aware Application; http://www.cl.cam.ac.uk/research/dtg/lce-pub/public/files/tr.2002.2.pdf }
+
+\bibitem{bat2}
+\texttt{ Andy Ward, Alan Jones and Andy Hopper; A New Location Technique for the Active Office; http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.49.1301&rep=rep1&type=pdf}
+
+\bibitem{radar}
+\texttt{Microsoft research: radar; https://www.microsoft.com/en-us/research/project/radar/}
+
+\bibitem{radar1}
+\texttt{P. Bahl and V. N. Padmanabhan; RADAR: An in-building RF-based user
+location and tracking system; http://morse.colorado.edu/~timxb/5520/ho/BahlRADAR.pdf}
+
+\bibitem{cricket}
+\texttt{Nissanka Bodhi Priyantha; The Cricket Indoor Location System;http://ahvaz.ist.unomaha.edu/azad/temp/sac/05-priyantha-localization-sensor-networks-coverage-thesis-good.pdf}
+
+\bibitem{cricket1}
+\texttt{Nissanka B. Priyantha, Anit Chakraborty, and Hari Balakrishnan; The Cricket Location-Support System; http://www.cse.wustl.edu/~lu/cse521s/Papers/cricket.pdf}
+
 \end{thebibliography}

### 2017-03-05 paper.tex
ok
SpellChecked everything
updated for jnos

@@
 \usepackage{caption}
+\usepackage{url}
 \usepackage{subcaption}
 \usepackage[caption=false, ...]{subfig}
-
-Generic and extensible indoor location services
-
-Multimodal indoor location service/framework
 
@@
 
-Indoor positioning systems have greatly evolved in the past few years due to the great success of its counter-part, the global positioning system (GPS), in outdoor environments but failure to reproduce the same results in indoor environments.
+Outdoor positioning systems have greatly evolved in the past with the global positioning system (GPS) as its most important system.
 Since GPS is an outdoor position systems and is based on a network of satelite, when the required scenario for position tracking is inside a building, new constraints are presented onto the process such as the attenuation and reflection of eletromagnetic waves upon collision with building walls and obstacles \cite{survey1}.
-As such there was a need to find reliable indoor systems that by nature would already be able to heavily reduce the impact of some of the mentioned constraints.
+As such there was a need to find reliable indoor systems that by nature would reduce the impact of some of the mentioned constraints.
 
-In order to understand indoor position there is a need to understand the full scope of variables that come to surface when moving from outdoor to indoor.
-When developing an indoor system there is a need to make sure that it can tackle the chalenges such as small space dimension which reforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques}.
+In order to understand indoor position there is a need to understand the full scope of variables that comes to surface when moving from outdoor to indoor.
+When developing an indoor system there is a need to make sure that it can tackle the chalenges such as small space dimension which reforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques, survey3}.
 All of the previously mentioned affect the way electromagnetic waves propagate in an indoor environment leading to problems related to severe multipath and reflection on existent surfaces \cite{surveywireless}.
 Besides propagation challenges, there are energy consumption, accuracy and deployment costs that play a critical role in deciding the viability of a proposed indoor location technique.
+Another interesting challenges that also surfaces is the means to describe a location.
+When dealing with outdoor location systems, the provided location is always characterized by two values, latitude and longitude, since the existant environment, i.e.
+the planet Earth, is always the same and as such the coordinates system can be relative to it.
+In indoor systems, the surrounding environment can be of many shapes and as such different ways of representing the data are necessary.
+If one thinks about providing indoor location on an office, the precise location isn't as relevant as just knowing the general location, i.e.
+a building specific description of the location in the form of building, floor and room.
+On other environments such as supermarkets or even in the previous example, where the objective is to provide more a precise location description, a cartesian coordinate system (x,y) is required.
 
-With the evolution of mobile devices there has been a sizeable number of different technologies that can possibly be used in indoor locationing \cite{surveywireless,survey2,survey1} such as GPS-based technologies, using high sensitivity antenas to overcome GPS's indoor issues, RFID , Wireless LAN and Bluetooth among others, allowing even for hibrid systems which make use of more than one of the technologies mentioned above.
+With the evolution of mobile devices there has been a sizeable number of different technologies that can possibly be used in indoor locationing \cite{surveythesis,survey2} such as GPS-based technologies, using high sensitivity antenas to overcome GPS's indoor issues, RFID , Wireless LAN and Bluetooth among others, allowing even for hibrid systems which make use of more than one of the technologies mentioned above.
 
-Another important aspect besides the chosen technology is the location detection technique which are widely varied in terms of accuracy and complexity.
-The simplest available category would be proximity detection which has Cell of Origin (CoO) as its most famous technique, allowing for room-based detection by assuming that the mobile target is at the cell with the strongest receiving power.
-This technique is widely used by system using RFID, Bluetooth and Infrared.
-The next category would be triangulation which uses the geometric properties of the triangle to obtain a position.
-The method utilized to obtain the values used in the position's calculation can be angle-based, which have implementation costs are their worst limitation; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI) as a metric, altough it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position utilizing the last determined position and incrementing based on estimate of the device's speed.
+%% ADD Generic Problems (i.e. using smartphone)
+%  3 vectors: technology (beacons) , loc description , loc calculation
+
+We are now in an era dominated by smartphones and as such they have became the central piece of several indoor location system.
+Smartphone's evolution allowed for developing system which rely on its sensors and processing capacities to present a solution which isn't dependant on specific hardware.
+With such a dependancy, the first condition that is imposed onto the systems is the compatibility with smartphones, i.e.
+the required sensors needs to exist on the generic hardware of smartphones.
+Once this barrier is surpassed, these generic system's are immediately faced with three fundamental questions which will define the architecture of the system and which can be seen on figure \ref{fig:choices}.
+The first question, Technology, defines the technology which is to be used in conjunction with the smartphone and consequently the way that location data is to be collected.
+There is a wide range of possibilities for this choice, be it BLE beacons, Wi-Fi Access points, LED lamps or just the microphone for sound collection, and each has an impact on the way that the system functions and on its performance.
+The second question is the location algorithm, which fundamentally depends on the target requirements of the system, if it's required to provide accurate location of a user or if a more descriptive location, such as the room in which the user is located, is enough.
+The third and last question is about which way will the location be described, with the existant possibilities having previously been presented.
+
+\begin{figure}[htp]
+	\centering
+		\includegraphics[width=0.5\linewidth]{figures/vectors.png}
+	\caption[]{Three fundamental choices in Indoor systems}
+	\label{fig:choices}
+\end{figure}
 
@@
 This high number of variations has led to an increase on the number of existing surveys which attemp to gather make a collection of what has been achieved in the field.
-Looking at this opportunity I saw the possibility of attempting to create a generic platform capable of being utilized by any existing indoor location system.
-In order to implement and test the solution there was a need to decide on which technology to be used and as such I opted by the bluetooth low energy.
+Looking at this opportunity one saw the possibility of attempting to create a generic platform capable of being utilized by any existing indoor location system.
+In order to implement and test the solution there was a need to decide on which technology to be used and as such one opted by the bluetooth low energy.
 
-In this paper is structured in a way that in \ref{sec:related} an overview of Bluetooth low energy (BLE) and existent projects and their associated architecture are presented while shedding light on the benefits and limitations of the chosen technologies.
+% Describe generic solution
+The Idealized generic architecture , which can be seen on figure \ref{fig:generic}, was created with the intention of providing a common ground for any systems which planned to have an indoor location based on a smartphone, which allows interoperability among those who make use of it, while presenting a way to structure the system so that the impact on the smartphone is reduced.
+If one analyses it according to the three points presented in figure \ref{fig:choices}, one can say that by having dedicated serves for the location computation and the location description instead of having it locally on the smartphone, interoperability is highly improved.
+This setup allows for systems to implement whichever algorithm they intent for as long as the server is capable of distinguishing which algorithm is to be used, responsability which is passed onto the smartphone.
+In scenarios of interoperability, the smartphone should be able to use any of its sensors to capture data, dealing with the technology component, and tag them before sending it for computation on the server.
+The location description is contained on the map server and it is up to the system to choose which one to utilise, be it self-created or owned by a third party such as google maps or openstreet maps.
+For as long as the smartphone is aware of the type of location that he is to receive, interoperability should be achieved.
+
+\begin{figure}
+	\centering
+		\includegraphics[width=1\linewidth]{figures/generic.png}
+	\caption[Generic System's Architecture]{Generic System's Architecture}
+	\label{fig:generic}
+\end{figure}
+
+% Review
+
+This paper is structured in a way that in \ref{sec:related} an overview of Bluetooth low energy (BLE) and existent projects and their associated architecture are presented while shedding light on the benefits and limitations of the chosen technologies.
 Section \ref{sec:architecture} overviews the architecture of the proposed generic system, while section \ref{sec:structure} presents the finalized system by analyzing the utilized technologies.
@@
 
-Indoor location systems is a field that has only been researched for a few couple of years but from the very begining until now there has been a very large amount of changes.
-In order to understand the progresses made there is a need to dwell into the systems which functioned as bases for progress.
+%into
+In this chapter one will first analyse the systems that layed the foundations of indoor location systems in section \ref{subsec:dedicated}, in order to study systems that were fully dependant on having specific hardware for obtaining user location.
+Section \ref{subsec:gentech} will present the technlogies that are compatible with modern day smartphones, analyse its pros and cons and mention existant projects based on them.
+The following section, \ref{subsec:technique} will describe the existent ways of obtaining the data that is required for infering a user's location as well as the techniques that make use of them.
+To conclude, on section \ref{subsec:desciption}, one will present the existent ways of describing locations through system that make use of them.
 
-In 1992 the Active system \cite{badge} was presented as an infrared solution which was capable of provided room-based position tracking, due to  infrared signals being unable to to travel through walls, and whose's information was then stored in a centralized location database \cite{badge1}.
-This system imposed users to carry with them an ID card equiped with an IR LED which emitted a signal each fifteen seconds.
+\subsection{Dedicated Technologies}
+\label{subsec:dedicated}
+
+In 1992 the Active badge system \cite{badge} was presented as an infrared solution which was capable of provided room-based position tracking, due to infrared signals being unable to travel through walls.
+This system imposed users to carry with them an ID card equiped with an IR LED which emitted a signal each fifteen seconds \cite{badge1, badge2}.
 The long signal frequency was due to the chance of collision of signals from multiple users in the same location while also having in mind energy cost which need to be kept as low as possible.
@@
 The first one, Bat Ultrasonic Location System, was a system capable of tracking various objects, each with a small wireless device called bat \cite{bat1, bat2}.
-A bat consisted of a radio transmitter, each had a unique identifier and was ment to be carried by personnel or attached to objects.
-In order to locate a bat, there was a base station which periodically transmitted a signal with a single identifier which caused the corresponding bat to emmit an ultrasound pulse.
-The same pulse was then captured by an existing grid of receivers, which were placed on the ceiling of the areas to be covered, which recorded the time of arrival of the ultrasound.
-Since the receivers are also capable of receiving the initial base station's message they are capable of computing the ultrasound's time of flight, which is then forwarded to the base station for further computation.
-Since the speed of sound in the air is known, the system was capable of converting the time of flight of the ultrasound and convert it into bat-receiver distance.
-In order to obtain the bat's location, if at least three receivers had caught the bat's signal, it was possible to utilize multilateration to obtain its 3D position.
-Due to the nature of the system, only one bat can be localised at a time, as such the process was diveded into time slots which garanteed that each time slot had enough time to obtain a bat's position and for its emitted signals to die out.
+A bat consisted of a radio transmitter with a unique identifier and was ment to be carried by personnel or attached to objects.
+For locating a bat, the base station periodically transmitted a signal with a single ID which caused the corresponding bat to emmit an ultrasound pulse.
+This pulse was then captured by an existing grid of receivers, which were placed on the ceiling of the areas to be covered, which recorded the time of arrival of the ultrasound.
+Since the speed of sound in the air is known, the system was capable of converting the time of flight of the ultrasound into distances, in order to obtain the bat's location.
 
 The Radar system is a RF-based system to locate and track users inside buildings.
-The system is devided into two phases, an offline phase used for data collection and a real-time phase where a location is obtained.
+The system is devided into two phases, an offline phase used for data collection and a real-time phase where a location is obtained \cite{radar1}.
 For data collection the mobile application and the base stations are synchronized, in order to be able to obtain time stamps and then the mobile user periodically emits UDP packets to the base stations (BS) .
-Each BS records the RF signal strength measurement together with the time stamp.
-This information gathering happens in both offline and real-time phases.
-During only the offline phase, while the mobile user emits packets, it also indicates his position on the map through (x,y) coordinates together with the timestamp.
+Each BS records the RF signal strength measurement together with the time stamp, in order to create a fingerprint of the map.
 The real-time phase utilises the real-time user's packet associated RF signal strength, which is obtained at each bs and later on forwarded to the central computer where the computation is made, and examinates it in order to find the best fit for the current transmitter position.
 
-The last remarkable project, Cricket, was introduced in 2005 and managed to tackled some of the problems presented in the last mentioned systems \cite{cricket1}.
+The last remarkable project, Cricket, managed to tackled some of the problems presented in the last mentioned systems \cite{cricket1}.
 Each node in the cricket system is composed of a RF transceiver and hardware capable of generating and receiving ultrasonic signals.
@@
 
-TODO recent stuff
+\subsection{Generic Technologies}
+\label{subsec:gentech}
+
+Wi-Fi is a technology that can be used to estimate the location of a mobile user that resides inside the network.
+Nowadays Wi-Fi positioning systems have become the most widespread approach for indoor location systems since Wi-FI access points are readily available in many indoor environments and since all smartphones are equiped with a Wi-Fi antenna they can be located without the need of installing extra software or manipulating the hardware.
+One issue of Wi-Fi signals is that they suffer attenuation from static environment such as walls and movement of furniture and doors.
+The most widely location technique used is the radio signal strength indicator (RSSI), which suffers from severe multi-path effects leading to propagation model failures and as such inaccuracy in distance measurement.
+With these problems in mind a technique called RSSI-based fingerprinting is often used in order to improve performance, the same technique used in the Radar system.
+
+Infrared (IR) systems are one of the most common position system that utilise wireless technology that has been used to track objects or people.
+IR wavelengths are invisible to the human eye under most circumstances, making this technology less intrusive than those which are visible.
+This technology is completely available in all smartphones since it makes use of the camera for data input, requiring only line-of-sight communication between receiver and transmitter, preferably without interference from strong light sources.
+One example is Epsilon \cite{epsilon}, which makes use of Light-emitting Diode (LED) and light sensors.
+Epsilon presents a solution which is easy to use in today's world due to the availability of LED lighting in indoor environments, with the only requirement being the adaption of the existing infrastructure to the required by the system.
+In order for the system to work, it required the mobile user to carry a device with an incorporated light sensor, which is the case of pretty much any existent smartphone.
 
 Bluetooth is a wireless technology that was created in 1994 with the objective of replacing cables connecting fixed or portable devices.
-At this point in time Bluetooth Special Interest Group is in charge of developing and managing this technology characterized by its robustness, low energy consuption and low cost.
-The Bluetooth Low Energy protocol was introduced with the Bluetooth Core Specification version 4 (also called Bluetooth Smart) circa 2010 alongside two other protocols.
-Out of the three, BLE standed out for its lower power consuption, lower complexity and lower cost, while allowing for  device discovery, connection establishment and connection mechanisms.
-
-The BLE radio operates at the 2.4GHz band and employs a frequency hopping transceiver to combat interference and fading.
-It also employs two multiple access schemes: FDMA used to separate the 40 available physical channels, 37 of them are used as data channels and the remaining as advertising channels and TDMA in a polling scheme that is used when one device transmits a packet at a predetermined time and a corresponding device responds with a packet after a predetermined interval.
+The Bluetooth Low Energy protocol was introduced with the Bluetooth Core Specification version 4 (also called Bluetooth Smart) \cite{BLECore} and it standed out for its lower power consuption, lower complexity and lower cost, while allowing for  device discovery, connection establishment and connection mechanisms.
 
@@
 
+\subsection{Location Calculation techniques}
+\label{subsec:tecnique}
+
+The location of a user can be obtained through several different techniques, each possibility leads to different levels of accuracy and hardware and computacional costs.
+The first technique would be proximity detection which has Cell of Origin (CoO) as its most famous technique and allows for room-based detection by assuming that the mobile target is at the cell with the strongest receiving power.
+This technique is widely used by system using RFID, Bluetooth and Infrared and has low deployment costs as the number of beacons required is small, at least one per room.
+
+The following technique would be triangulation which uses the geometric properties of the triangle to obtain a position.
+The method utilized to obtain the values used in the position's calculation can be angle-based, which have implementation costs are their worst limitation; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI) as a metric, altough it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position utilizing the last determined position and incrementing based on estimate of the device's speed.
+Indoor location systems is a field that has only been researched for a few couple of years but from the very begining until now there has been a very large amount of changes.
+In order to understand the progresses made there is a need to dwell into the systems which functioned as bases for progress.
+
+\subsection{Location Description}
+\label{subsec:description}
+
+%Descriptions
+
+% Meridian
+
 \section{Architecture}
@@
 The generic system's architecture is presented in figure ~\ref{fig:generic} and it's divided in 4 parts: beacon, location server, map server and smartphone application.
-
-\begin{figure}
-	\centering
-		\includegraphics[width=1\linewidth]{figures/generic.png}
-	\caption[Generic System's Architecture]{Generic System's Architecture}
-	\label{fig:generic}
-\end{figure}
 
@@
 
-- Describe the Database ( insert image example of database with a few entries?? )
+%- Describe the Database ( insert image example of database with a few entries?? )
 
-- Mention capability to Insert aditional info, at the moment it displays geocoding on pop up menu in figure ~\ref{fig:AppMenu}.
+%- Mention capability to Insert aditional info, at the moment it displays geocoding on pop up menu in figure ~\ref{fig:AppMenu}.
 
@@
 The voting system uses an exponential function in order to attribute a weight to each server.
-INSERT FUNCTION AND EXPLAINATION.
+%INSERT FUNCTION AND EXPLAINATION.
 
@@
 
-\begin{thebibliography}{9}
-\bibitem{surveywireless}
- \texttt{Hui Liu; Survey of Wireless Indoor Positioning Techniques and Systems http://www.pitt.edu/~dtipper/2011/Survey1.pdf}
-
-\bibitem{reviewtechniques}
- \texttt{Zahid Farid, Rosdiadee Nordin, and Mahamod Ismail; Recent Advances in Wireless Indoor Localization Techniques and System}
-
-\bibitem{survey2}
- \texttt{Hakan Koyuncu, Shuang Hua Yang; A Survey of Indoor Positioning and Object Locating Systems; http://paper.ijcsns.org/07_book/201005/20100518.pdf}
-
-\bibitem{survey1}
- \texttt{Jiang Xiao, Zimu Zhou,Youwen Yi and Lionel M. Ni; A Survey on Wireless Indoor Localization from the Device Perspective; http://www.tik.ee.ethz.ch/~zzhou/paper/csur16-xiao.pdf}
-
-\bibitem{survey3}
-\texttt{Klaithem Nuaimi and Hesham Kamel; A survey of indoor positioning systems and algorithms; https://www.researchgate.net/publication/224242434_A_survey_of_indoor_positioning_systems_and_algorithms}
-
-\bibitem{blecore}
-\texttt{ }
-
-\bibitem{ibeacon}
-\texttt{ibeacon insider: what's ibeacon; http://www.ibeacon.com/what-is-ibeacon-a-guide-to-beacons/}
-
-\bibitem{ibeacon1}
-\texttt{ Jingjing Yang, Zhihui Wang and Xiao Zhang ;An iBeacon-based Indoor Positioning Systems for Hospitals; http://www.sersc.org/journals/IJSH/vol9_no7_2015/16.pdf}
-
-\bibitem{bleacc}
-\texttt{ R. Faragher and R. Harle; An Analysis of the Accuracy of Bluetooth Low Energy for Indoor Positioning Applications; http://www.cl.cam.ac.uk/~rmf25/papers/BLE.pdf}
-
-\bibitem{lifemap}
-\texttt{Yohan Chon and Hojung Cha; LifeMap: A SmartphoneBased Context Provider for Location-Based Services; https://pdfs.semanticscholar.org/cb50/e7344494f8b8fdfeba160f814cbfe3ce3356.pdf}
-
-\bibitem{zonith}
-\texttt{http://media.teldio.com/pdfs/Teldio_Indoor}
-
-\bibitem{badge}
-\texttt{The Active Badge System; http://www.cl.cam.ac.uk/research/dtg/attarchive/ab.html}
-
-\bibitem{badge1}
-\texttt{WANT, R., HOPPER, A., FALCAO, V., AND GIBBONS, J.;The Active Badge Location System;  http://www.cl.cam.ac.uk/research/dtg/www/publications/public/files/tr.92.1.pdf}
-
-\bibitem{bat}
-\texttt{ The Bat Ultrasonic Location System ;http://www.cl.cam.ac.uk/research/dtg/attarchive/bat/ }
-
-\bibitem{bat1}
-\texttt{Andy Harter, Andy Hopper, Pete Steggles, Andy Ward and Paul Webster;The Anatomy of a Context-Aware Application; http://www.cl.cam.ac.uk/research/dtg/lce-pub/public/files/tr.2002.2.pdf }
-
-\bibitem{bat2}
-\texttt{ Andy Ward, Alan Jones and Andy Hopper; A New Location Technique for the Active Office; http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.49.1301&rep=rep1&type=pdf}
-
-\bibitem{radar}
-\texttt{Microsoft research: radar; https://www.microsoft.com/en-us/research/project/radar/}
-
-\bibitem{radar1}
-\texttt{P. Bahl and V. N. Padmanabhan; RADAR: An in-building RF-based user
-location and tracking system; http://morse.colorado.edu/~timxb/5520/ho/BahlRADAR.pdf}
-
-\bibitem{cricket}
-\texttt{Nissanka Bodhi Priyantha; The Cricket Indoor Location System;http://ahvaz.ist.unomaha.edu/azad/temp/sac/05-priyantha-localization-sensor-networks-coverage-thesis-good.pdf}
-
-\bibitem{cricket1}
-\texttt{Nissanka B. Priyantha, Anit Chakraborty, and Hari Balakrishnan; The Cricket Location-Support System; http://www.cse.wustl.edu/~lu/cse521s/Papers/cricket.pdf}
-
-\end{thebibliography}
+\bibliographystyle{IEEEtran}
+\bibliography{paper}
 

### 2017-03-06 paper.tex
Prepared for new additions

@@
 
-This paper is structured in a way that in \ref{sec:related} an overview of Bluetooth low energy (BLE) and existent projects and their associated architecture are presented while shedding light on the benefits and limitations of the chosen technologies.
+This paper is structured in a way that in \ref{sec:related} an overview of existent indoor location systems, generic and dedicated, is given, followed by an analyses on the existent location techniques and the ways to decribe a location.
 Section \ref{sec:architecture} overviews the architecture of the proposed generic system, while section \ref{sec:structure} presents the finalized system by analyzing the utilized technologies.
@@
 This technique is widely used by system using RFID, Bluetooth and Infrared and has low deployment costs as the number of beacons required is small, at least one per room.
-
 The following technique would be triangulation which uses the geometric properties of the triangle to obtain a position.
-The method utilized to obtain the values used in the position's calculation can be angle-based, which have implementation costs are their worst limitation; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI) as a metric, altough it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position utilizing the last determined position and incrementing based on estimate of the device's speed.
-Indoor location systems is a field that has only been researched for a few couple of years but from the very begining until now there has been a very large amount of changes.
-In order to understand the progresses made there is a need to dwell into the systems which functioned as bases for progress.
+The metrics utilized to obtain the values used in the position's calculation can be angle-based, which have high implementation costs due to the required antenna's complexity; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI), altough it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position utilizing the last determined position and incrementing based on estimate of the device's speed.
 
@@
 
-%Descriptions
+In order to better understand the existent ways of describing locations it's better to look at existent solutions and the methods that they utilise.
+The most common method of describing location is through latitude and longitude, much like the gps does, as the biggest example that makes use of it is Google maps.
+Google maps, through its indoor maps platforms \cite{googlemaps} allows integration of indoor maps onto their google maps.
+Since google maps uses this type of location description, the indoor component of it follows the same routes.
+An indoor map on the platform is inserted onto its original geographical location and in the case of having multiple floors, it is possible to navigate through them.
 
-% Meridian
+Another method is to consider the indoor map as the reference and utilise a cartesion coordinate system, x and y, to represent a location.
+Meridian, an indoor location system developed by HP \cite{meridian} which presents functionalities such as route making and push notifications through its beacons, makes use of its platform for insertion of the indoor maps.
+As such there isn't the outdoor component that google indoor maps had,making it possible to use a (x,y) system relative to the building map.
+
+Although Meridian already provides some extra level of detail in comparison with google indoor maps, there is another example that should be mentioned called OpenStreetMaps.
+Although it started much like google, providing global data, due to its openness, many indoor projects surfaced such as OpenLevelUP \cite{openlevel}.
+This project makes use of OpenStreetMaps' current indoor tagging scheme \cite{opentagging}, which is intented to describe in the most complete and simple way a building.
+This makes available the number of floors, the type of elemets (room, wall, corridor, etc) and its connectors, like doors and escalators or elevators, allowing to understand clearly the map that is being analysed.
 
@@
 
-The beacon represents any form of hardware responsible for providing fixed reference points which are fundamental in calculating a user's position.
-A beacon needs to be be composed of two structural blocks: The Unique ID, which stores any information crucial to identifying a beacon, and the Hardware specific software component, which contains all the necessary software to accomplish communication with a certain technology.
+The environment with beacons represents any form of element responsible for providing fixed reference points which are fundamental in calculating a user's position.
+A beacon needs to be compatible with the usage of a smartphone, i.e.
+the device's sensors need to be able to capture the data.
+Using the previously presented technologies as examples, these "beacons" could be in fact BLE beacons , Wi-Fi access points, LED lamps or even something as simple as sound, with the data capturing being made through the smartphone's available antennas, camera or microphone.
+The beacon represents a position in the indoor environments and as such it needs to be uniquely identified, be it through its characteristics or associated information.
+In addition to these informations, each beacon needs to have an associated location server and provide information about it to the smartphone application.
 
-The location server is divided in three: the Communication block, which is responsible for creating, managing and terminating a connection with the mobile application; the location algorithm, architectural block responsible for applying a certain location algorithm to the data received from the application in conjuntion with the server's device information in order to obtain a concrete location; and the beacon manager which functions as a database where all the beacon associated with the server are stored.
+The location server is an external component where the information relative to the "beacons" utilised is stored.
+As such when a location request arrives from the smartphone, the received data can be translated onto physical locations which will later be used to compute the user's location.
+The algorithm used to obtain such location is entirely dependant on the system's creators.
+Once the location is obtain, it's sent back to the smartphone in addition to the location of the server where it should retrieve the associated map.
 
-The map server represents the architectural block responsible for providing the maps associated to the location of user obtained through the location service on the application.
-As such it requires a communication block which needs to be capable of communication with the map service of the application and takes care of creating, managing and terminating connections, and a map manager which stores the existing maps that are to be transfered to the application.
+The map server represents the architectural block responsible for providing the maps associated to the location of the user, location which was obtained through the location service on the application.
+The used map representation is up to the system creater for as long as it is in accordance with the remaining parts of the system.
+This means that the location provided by the location server needs to be something that is representable on the maps provided and capable of being comprehended by the smartphone application.
 
-The smartphone application is divided into three functional blocks: core application, location service and map service.
-The core application contains the core functionality of the application and is in charge of communicating with the existing services.
-Communication with the location service allows for requesting the user's location which is then forwarded to the map service, in order to obtain the associated map.
-The location service is composed by four blocks: the API component through which it communicates with the core of the application; the tag reader which is the functional block responsible for reading a certain type of tags (beacons) and as such its structure will depend on the type of beacon utilized by the service; the tag decoder, functional block responsible of utilizing the information obtained through the tag reader and process it in order to be forwarded to communication block; the communication block which takes care of the creating, managing and terminating connections with a location server.
+The smartphone application is the central piece of this architecture.
+It is responsible of discovering and communicating with the beacons existent in its environment through its sensors.
+This communication process allows the smartphone to obtain information of its surrounding and to obtain the address of the location server that is associated to the connected beacon, in order to later forward this same information to it.
+Upon communicating with the location server, the application expects to get back information relative to its positions as well as the address of the map server that is to be contacted.
+Through that address, the application is capable of obtaining the map relative to its positions and display the result to the user.
 
-It might be relevant to analyze some types of indoor location system that could diverge from this architecture, such as camera or Quick Response (QR) code-based systems.
-These systems would still follow the proposed architecture except for the beacon component which would not be existant and the location service's tag reader block wouldn't need any sort of communication since it would be responsible of utilizing the camera to capture either a fotograph or reading a QR core.
-
-\begin{figure}
-	\centering
-		\includegraphics[width=0.75\linewidth]{figures/altgeneric.png}
-	\caption[Adjustments to the application required for multiple indoor location systems]{Adjustments to the application required for multiple indoor location systems}
-	\label{fig:altgeneric}
-\end{figure}
-
-Although figure \ref{fig:generic} is representative of a system utilizing a single type of technology, the presented solution is scalable, allowing for insertion of additional indoor location systems onto a single architecture.
-The changes required for implementing such a scenario would be to add one extra layer of complexity on the android application as present on figure \ref{fig:altgeneric}.
-In this new architecture, the application's core would only communicate with the location manager and the latter would now be responsible of managing through all of the location services and request execution of whichever service would be more suited or available.
-The location manager would also be required to manage through the existant map services, which may or may not be in same number as the existant location services, and make sure that the data obtained from a certain location service is forwarded to the correct map service which must capable of processing it.
+The presented architecture was structured in a way that it is scalable and it allowed for interoperability.
+The idea behind such an architecture would be let different indoor building use their own version of the system but a user with the associated smartphone application would be able to transit between buildings and use the readily available beacons to always obtain its location.
+This would be achieved through a common architecture that allowed for self-contained implementations that could be access through a generic smartphone application.
+In order to analyse this assumption it is necessary to look back on the already analysed figure \ref{fig:choices}, which represents the main components of an indoor systems.
+This architecture is capable of externalizing from the smartphone component all the remaining.
+By making the smartphone the central piece of communication, the beacons are required to the ones to provide location data.
+By having a server dedicated to computing the location of an user, the system developers are allowed to use whichever algorithm they wish, since there is no dependency to the application.
+They are also the only ones responsible for having the beacon associated data, for giving those same beacons the location of its server and for computing a location that is in accordance with how they wish to store their maps on the maps server.
+The last remaining component is the location description which is completely passed onto the map server and whose only requirement is for it to be in accordance with its associated location server, in terms of location description.
 
@@
 	\centering
-		\includegraphics[width=1\linewidth]{figures/System.png}
+		\includegraphics[width=1\linewidth]{figures/implementation.png}
 	\caption[System's Architecture]{System's Architecture}
@@
 
-%- Describe the Database ( insert image example of database with a few entries?? )
-
-%- Mention capability to Insert aditional info, at the moment it displays geocoding on pop up menu in figure ~\ref{fig:AppMenu}.
-
 \subsection{ Application}
@@
 The first step indicates the gathering of information of the surroundings of the user's device.
-When a request is made to the provider, a scan for nearby bluetooth low energy devices is made which will put the smartphone in a state of listening for incoming ble advertisement packets for half a second ( NEED TO VERIFY VALUE AND JUSTIFY, THERE WAS A PAPER ON THIS TODO).
+When a request is made to the provider, a scan for nearby bluetooth low energy devices is made which will put the smartphone in a state of listening for incoming ble advertisement packets.
 During this scanning period, each time a device is found, the advertisement is registered in a list, which have a duplication prevention mechanism implemented.
@@
 The voting system uses an exponential function in order to attribute a weight to each server.
+
 %INSERT FUNCTION AND EXPLAINATION.
@@
 The information contained inside the received message, which was described in section ~\ref{subsec:server}, is then processed into the adequate class capable of storing a geographic location and the same is broadcasted from the mock location provider to its listener.
-
-One of the current limitations of the android API that deals with BLE is that the interface on the smartphone that is used to connect with a device has a fixed timeout time.
-MIGHT JUST NEED TO BE IMPROVED TODO
 
@@
 
-- Location Accuracy
-
 - Time for request completion
@@
 
-- Possible Improvement: Paralel ble connections; Improvements on lateration algorithm;  Rest depends on the Performance results
-
 - Conclusions

### 2017-03-06 paper.tex
New section order . 6/3

@@
 
-%% ADD Generic Problems (i.e. using smartphone)
-%  3 vectors: technology (beacons) , loc description , loc calculation
-
 We are now in an era dominated by smartphones and as such they have became the central piece of several indoor location system.
@@
 
-% Describe generic solution
 The Idealized generic architecture , which can be seen on figure \ref{fig:generic}, was created with the intention of providing a common ground for any systems which planned to have an indoor location based on a smartphone, which allows interoperability among those who make use of it, while presenting a way to structure the system so that the impact on the smartphone is reduced.
@@
 
-%into
 In this chapter one will first analyse the systems that layed the foundations of indoor location systems in section \ref{subsec:dedicated}, in order to study systems that were fully dependant on having specific hardware for obtaining user location.
@@
 
-\section{Architecture}
-\label{sec:architecture}
+\section{GEFILOC}
+\label{sec:system}
+
+% Resume
+
+\subsection{Deployment environment}
+\label{subsec:deployment}
+
+In order to understand how a system based on this generic architecture, figure \ref{fig:generic} has to be strutured, it is of relevance to analyse the system component's description and requirements that would allow it to be deployed.
+If one starts by analysing the beacons component, one can describe them as the reference points of the system which are in charge of providing nearby users with information relative to their surroundings.
+These beacons are always ready and available to the users and upon being contacted, they should inform the smartphone which has initiated the communication of the address of the beacon's associated location server.
+Since the beacon only communicates with the smartphone, it isn't required to have any other communication capacity other than that specific to its technology.
+The last requirement is that the beacon needs to be uniquely identified and have the address of its associated server, while being capable of transmitting that same information to the smartphone.
+
+The smartphone service is the component that is in charge of communicating with both the beacons and the location server.
+This service is a software component of the smartphone application which is available to being called when a location is requested.
+Once a location is requested by the application, the service should communicate with nearby beacons through one of its sensors.
+A service is associated with one and only one technology, as such if the application is to support more than one technlogy, multiple services should be implemented.
+This condition is relevant since the required sensor depends on the chosen technology and because it allows the service to quickly associate the data that it is collecting with the technology, information which must be delivered to the location server.
+Once the beacon data has been collected, the same data is forwarded to the beacon's associated server, together with a tag representing the technology of the beacons.
+Upon obtaining a location, the service should forward the data to the application, which is responsible of later communicating with the map server.
+In terms of requirements, the service needs to have Wi-Fi or at least 3G available for communicating with both the location and maps server, and access to the sensor required to communicate with the nearby beacon, for BLE it would be the bluetooth antenna or for QR it would be the camera.
+
+The Location server should have a database of all the devices associated to it, each with its exact location of the map.
+This location is dependent on the method for location description chosen, i.e.
+latitude/longitude or x/y.
+The server should also be able to take input from any of the supported technologies, depends on the beacons in the database, in order to output the user's location.
+From this description one can define the requirements of the location server as having in storage the beacons that are associated it as well as their location, being capable of communicating with the smartphone application and being in accordance with the map server, achieved through providing location description that is the same as that present in the maps server.
+
+The Map server is responsible for providing the application with the maps that it has requested.
+The only decision concerning this server is the type of location description that is to be used and consequently the only requirement is that it has to be in accordance with the information on the location server.
+
+\subsection{Architecture}
+\label{subsec:architecture}
 
@@
 
-\section{Implementation}
+\subsection{Implementation}
 \label{sec:struture}
 
-\subsection{Bluetooth Low energy}
+The implementation presented in this paper was created by utilizing the generic indoor location system presented in section \ref{sec:architecture} and applying it with bluetooth low energy.
+The system's architecture is presented in figure ~\ref{fig:implementation} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:app} alongside figures that show the functional prototype.
+For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
+
+\begin{figure}
+	\centering
+		\includegraphics[width=1\linewidth]{figures/implementation.png}
+	\caption[System's Architecture]{System's Architecture}
+	\label{fig:implementation}
+\end{figure}
+
+\subsubsection{Bluetooth Low energy}
 \label{ble}
@@
 
-The implementation presented in this paper was created by utilizing the generic indoor location system presented in section \ref{sec:architecture} and applying it with bluetooth low energy.
-The system's architecture is presented in figure ~\ref{fig:implementation} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:app} alongside figures that show the functional prototype.
-For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
-
-\begin{figure}
-	\centering
-		\includegraphics[width=1\linewidth]{figures/implementation.png}
-	\caption[System's Architecture]{System's Architecture}
-	\label{fig:implementation}
-\end{figure}
-
-\subsection{ BLE beacon}
+\subsubsection{ BLE beacon}
 \label{subsec:beacon}
@@
 
-\subsection{ Server}
+\subsubsection{ Smartphone service}
+\label{subsec:service}
+
+\subsubsection{ Server}
 \label{subsec:server}
@@
 
-\subsection{ Application}
-\label{subsec:app}
+\subsection{Evaluation}
+\label{sec:eval}
 
@@
 
-\section{Performance Analysis}
-\label{sec:performance}
+%% Qualitativamente 
+%% graficos e tal
 
@@
 - Data Requiremnts?
+
+%% Qualitativamente
+
+Although a complete implementation of the proposed architecture was achieved for BLE, one must indicate the required changes that would be necessary to make when change one or several parts of the system.
+The parts that can be changed are the three possibilities previously studied from figure \ref{fig:choices}.
+In case the technlogy is to be change, we will use QR code as an example, the affected areas are the beacons, the smartphone service and the location server.
+By changing the to QR code, the technology changes instantly as there is no more requirement for BLE beacons but instead there will be QR codes placed in fixed places.
+Each QR code would be required contain information about its associated server.
+The service would change the mechanism utilised to collect data, instead of listening for BLE beacons through its bluetooth antenna, it will open up the camera, so that the user can capture the QR code which would then be forwarded to the location server.
+The location server would have to be prepared for receiving the QR data and obtaining the location relative to that QR code.
+
+If the system were to be changed on the location algorithm component, this would only require changes on the location server.
+The algorithms connection has only two connections points with the system, the data that it receives and the location it outputs.
+As such, for as long as the new algorithm makes use of the same data and in the end, once it has concluded, it outputs one location of the same format, no other issue should surface.
+
+A modification on the location description utilised would have an impact on both the location and map server.
+With such a request, the map server is required to change completely the stored maps in order to support the new description method, while the changes on the location server are on the database, since the location of the beacons should change, and on the way the location algorithm processes these same locations.
 

### 2017-03-12 paper.tex
update changes

@@
 
-The implementation presented in this paper was created by utilizing the generic indoor location system presented in section \ref{sec:architecture} and applying it with bluetooth low energy.
-The system's architecture is presented in figure ~\ref{fig:implementation} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:app} alongside figures that show the functional prototype.
+The implementation presented in this paper was created by utilizing the generic indoor location system presented in section \ref{subsec:architecture} and applying it with bluetooth low energy.
+The system's architecture is presented in figure ~\ref{fig:implementation} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:service} alongside figures that show the functional prototype.
 For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
@@
 \subsubsection{Bluetooth Low energy}
-\label{ble}
+\label{subsec:ble}
 
@@
 \label{subsec:service}
-
-\subsubsection{ Server}
-\label{subsec:server}
-
-The webserver was implement in Python 3 programming language.
-The program implements a simple tcp server capable of receiving multiple request at the same time.
-Each request starts with information sent from an application which include a pair of MAC address and associated RSSI value for each ble device that the same application found.
-Afterwards the list of pairs is filtered in order to remove any existant devices that are not present in the server's database of devices.
-
-Each server has a database that includes only ble devices.
-An entry (description of a device) in this database is composed by the device's mac address, its longitude and latitude and its building, floor and room name.
-In addition to the database, a server when initiated can store additional location info such as the server's street, number, zipcode, city and country, allowing this information to be transmited to the client in order to offer an additional level of location description to the user.
-The whole location specific information can be visualised in figure ~\ref{fig:AppMenu}.
-
-Upon having filtered the initial list of pairs, the Cell of Origin (CoO) technique is applied by verifying which of the devices produced a stronger signal on the receptor.
-Upon obtaining the closest device an answer is sent to the application containing all of the information associated to the server and the selected device.
-
-\subsection{Evaluation}
-\label{sec:eval}
 
@@
 
+\subsubsection{ Server}
+\label{subsec:server}
+
+The webserver was implement in Python 3 programming language.
+The program implements a simple tcp server capable of receiving multiple request at the same time.
+Each request starts with information sent from an application which include a pair of MAC address and associated RSSI value for each ble device that the same application found.
+Afterwards the list of pairs is filtered in order to remove any existant devices that are not present in the server's database of devices.
+
+Each server has a database that includes only ble devices.
+An entry (description of a device) in this database is composed by the device's mac address, its longitude and latitude and its building, floor and room name.
+In addition to the database, a server when initiated can store additional location info such as the server's street, number, zipcode, city and country, allowing this information to be transmited to the client in order to offer an additional level of location description to the user.
+The whole location specific information can be visualised in figure ~\ref{fig:AppMenu}.
+
+Upon having filtered the initial list of pairs, the Cell of Origin (CoO) technique is applied by verifying which of the devices produced a stronger signal on the receptor.
+Upon obtaining the closest device an answer is sent to the application containing all of the information associated to the server and the selected device.
+
+\subsection{Evaluation}
+\label{sec:eval}
+
 The pop-up menu was implemented to demonstrate the capacity of providing additional information associated with each location, be it geo-location taxonomy as it is currently implemented or possibly a description of the located room, an hyperlink of some sort or any other type of data that someone implemented this system would like to provide to its users.

### 2017-03-21 paper.tex
Abstract

@@
 
-	Abstract
+Nowadays there is a great demand for positioning systems and while the outdoor perfomance is amazing through the use of GPS, indoor positioning systems have been studied in order to also achieve such results.
+In this paper one studies the actual state of indoor systems, through an analysis on the usable technologies and techniques, and makes a critical analysis through a modern prespective centered on location services through smartphones.
+Since there is a great number of different indoor location systems, one found a need to create an framework capable of hosting such systems in the form of a smartphone-centric architecture, where interoperability would attempt to be achieved through the externalization of the main components of indoor systems.
+As such GEFILOC, a generic framework for indoor location systems which would allow for system that made use of it, to be capable of operating together on the same device without additional requirements , is proposed and a Bluetooth low energy-based system is implemented.
+This implementation allows for analysing the energetic costs of the architecture and comprehending how to set up one or more systems for analysing the interoperability component.
 
@@
 
-%INSERT FUNCTION AND EXPLAINATION.
-
 The voting system was implemented with the objective providing a thin security layer by allowing multiple devices of the same server to overcome a single attacker's device which happened to be close to the user.

### 2017-04-15 paper.tex
Final V1

--- 2017-03-12 paper.tex
@@
 \title{{ \normalsize Master Thesis} \\ 
-	Bluetooth Low Energy Supported Indoor Location}
+GEFILOC: Generic Framework for Indoor Location Systems} 
 \author{ 
@@
 
-	Abstract
+Nowadays there is a great demand for positioning systems and while the outdoor performance is amazing through the use of GPS, indoor positioning systems have been studied in order to also achieve such results.
+In this paper one studies the actual state of indoor systems, through an analysis on the usable technologies and techniques.
+A critical analysis is conducted through a modern perspective centered on location services through smartphones.
+Since there is a great number of different indoor location systems, one found a need to create a framework capable of hosting such systems in the form of a smartphone-centric architecture.
+This generic architecture attempted to achieve  interoperability  through the isolation of the main components of indoor systems.
+As such GEFILOC, a generic framework for indoor location systems, was created.
+GEFILOC would allow for systems that made use of it, to be capable of operating together on the same device without additional requirements.
+In order to test the proposed framework, a Bluetooth low energy-based system was implemented.
+This implementation allowed for analysing the energetic costs of the architecture and comprehending how to set up one or more systems for analysing the interoperability component.
 
@@
 Outdoor positioning systems have greatly evolved in the past with the global positioning system (GPS) as its most important system.
-Since GPS is an outdoor position systems and is based on a network of satelite, when the required scenario for position tracking is inside a building, new constraints are presented onto the process such as the attenuation and reflection of eletromagnetic waves upon collision with building walls and obstacles \cite{survey1}.
+Since GPS is an outdoor position systems based on a network of satellite, when it's used for indoor location, it isn't capable of having the same performance.
+This is caused by the constraints introduced by indoor environments such as attenuation and reflection of electromagnetic waves upon collision with building walls and obstacles \cite{survey1}.
 As such there was a need to find reliable indoor systems that by nature would reduce the impact of some of the mentioned constraints.
@@
 In order to understand indoor position there is a need to understand the full scope of variables that comes to surface when moving from outdoor to indoor.
-When developing an indoor system there is a need to make sure that it can tackle the chalenges such as small space dimension which reforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques, survey3}.
+When developing an indoor system there is a need to make sure that it can tackle the challenges such as small space dimension which reinforce the need for higher precision, a higher probability of non existent line of sight, influence of obstacles such as walls, furniture, moveable objects such as doors and human beings\cite{reviewtechniques, survey3}.
 All of the previously mentioned affect the way electromagnetic waves propagate in an indoor environment leading to problems related to severe multipath and reflection on existent surfaces \cite{surveywireless}.
@@
 Another interesting challenges that also surfaces is the means to describe a location.
-When dealing with outdoor location systems, the provided location is always characterized by two values, latitude and longitude, since the existant environment, i.e.
+When dealing with outdoor location systems, the provided location is always characterised by two values, latitude and longitude, since the existent environment, i.e.
 the planet Earth, is always the same and as such the coordinates system can be relative to it.
-In indoor systems, the surrounding environment can be of many shapes and as such different ways of representing the data are necessary.
+In indoor systems, there are many different types of surrounding environments and as such different ways of representing the data are necessary.
 If one thinks about providing indoor location on an office, the precise location isn't as relevant as just knowing the general location, i.e.
@@
 
-With the evolution of mobile devices there has been a sizeable number of different technologies that can possibly be used in indoor locationing \cite{surveythesis,survey2} such as GPS-based technologies, using high sensitivity antenas to overcome GPS's indoor issues, RFID , Wireless LAN and Bluetooth among others, allowing even for hibrid systems which make use of more than one of the technologies mentioned above.
+With the evolution of mobile devices there has been a sizeable number of different technologies that can possibly be used in indoor location \cite{surveythesis,survey2} such as GPS-based technologies, using high sensitivity antenas to overcome GPS's indoor issues, RFID , Wireless LAN and Bluetooth among others, allowing even for hybrid systems which make use of more than one of the technologies mentioned above.
 
 We are now in an era dominated by smartphones and as such they have became the central piece of several indoor location system.
-Smartphone's evolution allowed for developing system which rely on its sensors and processing capacities to present a solution which isn't dependant on specific hardware.
-With such a dependancy, the first condition that is imposed onto the systems is the compatibility with smartphones, i.e.
+Smartphone's evolution allowed for developing system which rely on its sensors and processing capacities to present a solution that isn't dependent on specific hardware.
+With such a dependency, the first condition that is imposed onto the systems is the compatibility with smartphones, i.e.
 the required sensors needs to exist on the generic hardware of smartphones.
 Once this barrier is surpassed, these generic system's are immediately faced with three fundamental questions which will define the architecture of the system and which can be seen on figure \ref{fig:choices}.
-The first question, Technology, defines the technology which is to be used in conjunction with the smartphone and consequently the way that location data is to be collected.
+The first question, Technology, defines the technology that is to be used in conjunction with the smartphone and consequently the way that location data is to be collected.
 There is a wide range of possibilities for this choice, be it BLE beacons, Wi-Fi Access points, LED lamps or just the microphone for sound collection, and each has an impact on the way that the system functions and on its performance.
 The second question is the location algorithm, which fundamentally depends on the target requirements of the system, if it's required to provide accurate location of a user or if a more descriptive location, such as the room in which the user is located, is enough.
-The third and last question is about which way will the location be described, with the existant possibilities having previously been presented.
+The third and last question is about which way will the location be described, with the existent possibilities having previously been presented.
 
@@
 
-When taking into consideration every possible technology and existant algorithm that are capable of being applied on indoor location, we find ourselves with a wide variaty of solutions.
-This high number of variations has led to an increase on the number of existing surveys which attemp to gather make a collection of what has been achieved in the field.
-Looking at this opportunity one saw the possibility of attempting to create a generic platform capable of being utilized by any existing indoor location system.
+When taking into consideration every possible technology and existent algorithm that are capable of being applied on indoor location, we find ourselves with a wide variety of solutions.
+This high number of variations has led to an increase in the number of existing surveys which attempt to gather make a collection of what has been achieved in the field.
+Looking at this opportunity one saw the possibility of attempting to create a generic platform capable of being used by any existing indoor location system.
 In order to implement and test the solution there was a need to decide on which technology to be used and as such one opted by the bluetooth low energy.
 
-The Idealized generic architecture , which can be seen on figure \ref{fig:generic}, was created with the intention of providing a common ground for any systems which planned to have an indoor location based on a smartphone, which allows interoperability among those who make use of it, while presenting a way to structure the system so that the impact on the smartphone is reduced.
-If one analyses it according to the three points presented in figure \ref{fig:choices}, one can say that by having dedicated serves for the location computation and the location description instead of having it locally on the smartphone, interoperability is highly improved.
-This setup allows for systems to implement whichever algorithm they intent for as long as the server is capable of distinguishing which algorithm is to be used, responsability which is passed onto the smartphone.
+The idealised generic architecture, which can be seen on figure \ref{fig:generic}, was created with the intention of providing a common ground for any systems which planned to have an indoor location based on a smartphone.
+It would allow for interoperability among those who make use of it, while presenting a way to structure the system so that the impact on the smartphone is reduced.
+If one analyses it according to the three points presented in figure \ref{fig:choices}, one can say that by having dedicated servers for the location calculation and the location description instead of having it locally on the smartphone, interoperability is highly improved.
+This setup allows for systems to implement whichever algorithm they intent, for as long as the server is capable of distinguishing which algorithm is to be used.
 In scenarios of interoperability, the smartphone should be able to use any of its sensors to capture data, dealing with the technology component, and tag them before sending it for computation on the server.
-The location description is contained on the map server and it is up to the system to choose which one to utilise, be it self-created or owned by a third party such as google maps or openstreet maps.
+The location description is contained on the map server and it is up to the system to choose which one to use, be it self-created or owned by a third party such as google maps or openstreet maps.
 For as long as the smartphone is aware of the type of location that he is to receive, interoperability should be achieved.
@@
 
-% Review
-
-This paper is structured in a way that in \ref{sec:related} an overview of existent indoor location systems, generic and dedicated, is given, followed by an analyses on the existent location techniques and the ways to decribe a location.
-Section \ref{sec:architecture} overviews the architecture of the proposed generic system, while section \ref{sec:structure} presents the finalized system by analyzing the utilized technologies.
-Section \ref{sec:performance} analyzes the different aspects of the presented solution in terms of energetic efficiency, accuracy and response timings, finalizing in section \ref{sec:future} by overviewing what could be carried out in order to further develop the existing work and by concluding the paper.
+This paper is structured in a way that in section \ref{sec:related} an overview of existent indoor location systems, generic and dedicated, is given, followed by an analyses on the existent location techniques and the ways to describe a location.
+Section \ref{sec:system} is entirely dedicated to the GEFILOC.
+The section starts by describing the framework's requirements in subsection \ref{subsec:deployment}, followed by an overview of the architecture in subsection \ref{subsec:architecture}.
+Subsection \ref{sec:struture} presents the Bluetooth low energy implemented system, while subsection \ref{sec:eval} presents a critical analysis on the implemented system's energetic consumption.
 
@@
 
-In this chapter one will first analyse the systems that layed the foundations of indoor location systems in section \ref{subsec:dedicated}, in order to study systems that were fully dependant on having specific hardware for obtaining user location.
-Section \ref{subsec:gentech} will present the technlogies that are compatible with modern day smartphones, analyse its pros and cons and mention existant projects based on them.
-The following section, \ref{subsec:technique} will describe the existent ways of obtaining the data that is required for infering a user's location as well as the techniques that make use of them.
-To conclude, on section \ref{subsec:desciption}, one will present the existent ways of describing locations through system that make use of them.
+In this chapter one will first analyse the systems that laid the foundations of indoor location systems in section \ref{subsec:dedicated}, in order to study systems that were fully dependent on having specific hardware for obtaining user location.
+Section \ref{subsec:gentech} will present the technologies that are compatible with modern day smartphones, analyse its pros and cons and mention existent projects based on them.
+The following section, \ref{subsec:tecnique} will describe the existent ways of obtaining the data that is required for inferring a user's location as well as the techniques that make use of them.
+To conclude, on section \ref{subsec:description}, one will present the existent ways of describing locations through system that make use of them.
 
@@
 In 1992 the Active badge system \cite{badge} was presented as an infrared solution which was capable of provided room-based position tracking, due to infrared signals being unable to travel through walls.
-This system imposed users to carry with them an ID card equiped with an IR LED which emitted a signal each fifteen seconds \cite{badge1, badge2}.
-The long signal frequency was due to the chance of collision of signals from multiple users in the same location while also having in mind energy cost which need to be kept as low as possible.
+This system imposed users to carry with them an ID card equipped with an IR LED which emitted a signal each fifteen seconds \cite{badge1, badge2}.
+The long signal frequency was due to the chance of collision of signals from multiple users in the same location while also keeping in mind energy cost, which need to be kept as low as possible.
 One of the downsides was the decrease in accuracy as the user's location is only known at best to a 15-second window.
-The user's location was obtained  through the implementation of a network of sensors which act as receivers and then forward the obtained information to the master station which would take care of polling the sensors, processing data and then making it available to users.
+The user's location was obtained  through the implementation of a network of sensors, which acted as receivers and forwarded the obtained information to the master station.
+The master station was responsible for polling the sensors, processing data and then making it available to users.
 One of the biggest issues that were raised with this system was the privacy concerns raised by the clients due to its object tracking nature.
 
-At the beggining of the twentieth century, two remarkable projects were presented, Bat system \cite{bat} and RADAR\cite{radar}.
+At the beginning of the twentieth century, two remarkable projects were presented, Bat system \cite{bat} and RADAR\cite{radar}.
 The first one, Bat Ultrasonic Location System, was a system capable of tracking various objects, each with a small wireless device called bat \cite{bat1, bat2}.
-A bat consisted of a radio transmitter with a unique identifier and was ment to be carried by personnel or attached to objects.
-For locating a bat, the base station periodically transmitted a signal with a single ID which caused the corresponding bat to emmit an ultrasound pulse.
+A bat consisted of a radio transmitter with a unique identifier and was meant to be carried by personnel or attached to objects.
+For locating a bat, the base station periodically transmitted a signal with a single ID which caused the corresponding bat to emit an ultrasound pulse.
 This pulse was then captured by an existing grid of receivers, which were placed on the ceiling of the areas to be covered, which recorded the time of arrival of the ultrasound.
@@
 The Radar system is a RF-based system to locate and track users inside buildings.
-The system is devided into two phases, an offline phase used for data collection and a real-time phase where a location is obtained \cite{radar1}.
-For data collection the mobile application and the base stations are synchronized, in order to be able to obtain time stamps and then the mobile user periodically emits UDP packets to the base stations (BS) .
-Each BS records the RF signal strength measurement together with the time stamp, in order to create a fingerprint of the map.
-The real-time phase utilises the real-time user's packet associated RF signal strength, which is obtained at each bs and later on forwarded to the central computer where the computation is made, and examinates it in order to find the best fit for the current transmitter position.
+The system is divided into two phases, an offline phase used for data collection and a real-time phase where a location is obtained \cite{radar1}.
+For data collection the mobile application and the base stations are synchronised, in order to be able to obtain time stamps and then the mobile user periodically emits UDP packets to the base stations (BS) .
+Each BS recorded the RF signal strength measurement together with the time stamp, in order to create a fingerprint of the map.
+The real-time phase utilises the real-time user's packet associated RF signal strength, that is obtained at each base station.
+This information is then forwarded to the central computer where the computation is made.
 
@@
 Each listener utilizes the RF signal's beacon information alongside the obtain distances to beacons to compute their space position and orientation.
-Crickets Architecture allowed to solve the user pribacy, decentralized and ease of deployment issues there were present in the remaining projects \cite{cricket}.
+Cricket's architecture allowed to solve the user privacy, decentralised and ease of deployment issues there were present in the remaining projects \cite{cricket}.
 
@@
 Wi-Fi is a technology that can be used to estimate the location of a mobile user that resides inside the network.
-Nowadays Wi-Fi positioning systems have become the most widespread approach for indoor location systems since Wi-FI access points are readily available in many indoor environments and since all smartphones are equiped with a Wi-Fi antenna they can be located without the need of installing extra software or manipulating the hardware.
+Nowadays Wi-Fi positioning systems have became the most widespread approach for indoor location systems since its access points are readily available in many indoor environments and since all smartphones are equipped with a Wi-Fi antenna they can be located without the need of installing extra software or manipulating the hardware.
 One issue of Wi-Fi signals is that they suffer attenuation from static environment such as walls and movement of furniture and doors.
@@
 
-Infrared (IR) systems are one of the most common position system that utilise wireless technology that has been used to track objects or people.
+Infrared (IR) systems are mostly used for tracking objects or people.
 IR wavelengths are invisible to the human eye under most circumstances, making this technology less intrusive than those which are visible.
 This technology is completely available in all smartphones since it makes use of the camera for data input, requiring only line-of-sight communication between receiver and transmitter, preferably without interference from strong light sources.
-One example is Epsilon \cite{epsilon}, which makes use of Light-emitting Diode (LED) and light sensors.
-Epsilon presents a solution which is easy to use in today's world due to the availability of LED lighting in indoor environments, with the only requirement being the adaption of the existing infrastructure to the required by the system.
+One example of an IR system is Epsilon \cite{epsilon}, which makes use of Light-emitting Diode (LED) and light sensors.
+Epsilon presents a solution which is easy to use in today's world due to the availability of LED lighting in indoor environments, with the only requirement being the adaptation of the existing infrastructure to the required by the system.
 In order for the system to work, it required the mobile user to carry a device with an incorporated light sensor, which is the case of pretty much any existent smartphone.
@@
 Bluetooth is a wireless technology that was created in 1994 with the objective of replacing cables connecting fixed or portable devices.
-The Bluetooth Low Energy protocol was introduced with the Bluetooth Core Specification version 4 (also called Bluetooth Smart) \cite{BLECore} and it standed out for its lower power consuption, lower complexity and lower cost, while allowing for  device discovery, connection establishment and connection mechanisms.
+The Bluetooth Low Energy protocol was introduced with the Bluetooth Core Specification version 4 (also called Bluetooth Smart) \cite{BLECore} and it stood out for its lower power consumption, lower complexity and lower cost, while allowing for  device discovery, connection establishment and connection mechanisms.
 
@@
 
-When looking at what's possible to achieve using the BLE technology there is the example of Apple's creation iBeacon \cite{ibeacon} which was presented in 2013 with the porpuse of implementing proximity sensing systems.
-The device is capable of playing on the broadcaster role and as such its objective is to send nearby compatible receivers certain information.
-Some examples of application are to track customers or trigger location-based actions on devices such as push notifications or checking in on social media, with pratical cases such as the usage of iBeacons by McDonalds to offer special offers to their customers in their fast-food stores.
-An indoor location system utilizing this technology was presented by Jingjing Yang et al \cite{ibeacon1}, where these devices were used to indicate a pacient of his whereabouts through the proximity sensing proprieties and this information was later transfered over to a server in order to give clients a variaty of different services, from pacient counting, to nearby department's information and offer indoor guidance to the nearest available bed.
+When looking at what's possible to achieve using the BLE technology there is the example of Apple's creation iBeacon \cite{ibeacon} which was presented in 2013 with the purpose of implementing proximity sensing systems.
+The device is capable of being a broadcaster, whose objective is to send information to nearby compatible receivers.
+The iBeacon can be used to track customers or trigger location-based actions on devices such as push notifications or checking in on social media.
+As an example, McDonalds used iBeacons to offer special offers to their customers in their fast-food stores.
+An indoor location system using this technology was presented by Jingjing Yang et al \cite{ibeacon1}, where these devices were used to indicate a patient of his whereabouts through the proximity sensing properties.
+The user's location was later transferred to a server in order to give clients a variety of different services, from patient counting, to nearby department's information and offer indoor guidance to the nearest available bed.
 
-When utilizing BLE for indoor location the usual metric used to calculate distances is the RSSI.
-This metric withint the context of bluetooth brings to surface several issues such as the fact that RSSI as a metric is very accurate only when the target is within a meter of the beacon, since the value decreases as the inverse of the square of the distance to the beacon .
-As such when developing solutions for indoor location that require system with high accuracy capable of tracking moving objects, the usage of RSSI can't be utilized without further work.
+When employing BLE for indoor location, the usual metric used to calculate distances is the RSSI.
+This metric within the context of Bluetooth brings to surface several issues such as the fact that RSSI as a metric is very accurate only when the target is within a meter of the beacon, since the value decreases as the inverse of the square of the distance to the beacon.
+As such when developing solutions for indoor location that require systems with high accuracy, capable of tracking moving objects, the usage of RSSI isn't viable without further work.
 Faragher et al \cite{bleacc} tackled one of the techniques used to improve BLE system's accuracy, fingerprinting, by verifying the effects caused by the device deployment density within the required location.
-This experiment also puts into evidence one of the downsides of the bluetooth technology being that its scalability is low, besides requiring higher density in order to increase accuracy, due to their low range any need to increase coverage leads to increased costs.
+This experiment also puts into evidence one of the downsides of the Bluetooth technology being that its scalability is low.
+Due to their low range, in order to increase accuracy, an higher density is required .
 
-Zonith \cite{zonith} introduced a bluetooth based location system with the objective of tracking the position of workers in dangerous environments.
-Any device registered in the zonith implemented network would be continuously tracked and accounted for in each of the system's functionalities such as, sounding an alarm whenever a lone worker doesn't move or responde within a time interval (Lone worker protection) or providing a quick an precise location of any worker that has requested for help.
-This system's installation requires planing of the best locations to place the beacons and number required of beacons in order to be able to provide enough courage and make sure the system provides the required quality.
+Zonith \cite{zonith} introduced a Bluetooth based location system with the objective of tracking the position of workers in dangerous environments.
+Any device registered in the zonith implemented network would be continuously tracked and accounted for in each of the system's functionalities.
+The system's functionalities involve sounding an alarm whenever a lone worker doesn't move or respond within a time interval (Lone worker protection) or providing a quick a precise location of any worker that has requested for help.
+This system's deployment requires a study on the best locations for placing the beacons and on the number of required beacons, so that it can provide enough coverage and assure the system's quality.
 
@@
 
-The location of a user can be obtained through several different techniques, each possibility leads to different levels of accuracy and hardware and computacional costs.
+The location of a user can be obtained through several different techniques, with each possibility leading to different levels of accuracy and hardware and computational costs. 
 The first technique would be proximity detection which has Cell of Origin (CoO) as its most famous technique and allows for room-based detection by assuming that the mobile target is at the cell with the strongest receiving power.
@@
 The following technique would be triangulation which uses the geometric properties of the triangle to obtain a position.
-The metrics utilized to obtain the values used in the position's calculation can be angle-based, which have high implementation costs due to the required antenna's complexity; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI), altough it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position utilizing the last determined position and incrementing based on estimate of the device's speed.
+The metrics used in the position's calculation can be angle-based, which have high implementation costs due to the required antenna's complexity; time-based, using Time-of-flight (ToF) or Round Trip Time (RTT) to obtain measures of distance from beacons; signal property based, using received signal strength indicator (RSSI), although it's only possible of using this metric with radio signals; dead reckoning, which calculates the actual position through the last determined position and an estimate of the device's speed.
 
@@
 
-In order to better understand the existent ways of describing locations it's better to look at existent solutions and the methods that they utilise.
-The most common method of describing location is through latitude and longitude, much like the gps does, as the biggest example that makes use of it is Google maps.
-Google maps, through its indoor maps platforms \cite{googlemaps} allows integration of indoor maps onto their google maps.
-Since google maps uses this type of location description, the indoor component of it follows the same routes.
-An indoor map on the platform is inserted onto its original geographical location and in the case of having multiple floors, it is possible to navigate through them.
+In order to better understand the existent ways of describing locations it's better to look at existent solutions and the methods that they employ.
+The most common method of describing location is through latitude and longitude, much like the GPS does, a method used by Google maps.
+Google maps, through its indoor maps platforms \cite{googlemaps} allows integration of indoor maps onto their maps.
+Since Google Maps uses this type of location description, its indoor component follows the same format.
+An indoor map on the platform is inserted into its original geographical location and in the case of having multiple floors, it is possible to navigate through them.
 
-Another method is to consider the indoor map as the reference and utilise a cartesion coordinate system, x and y, to represent a location.
-Meridian, an indoor location system developed by HP \cite{meridian} which presents functionalities such as route making and push notifications through its beacons, makes use of its platform for insertion of the indoor maps.
-As such there isn't the outdoor component that google indoor maps had,making it possible to use a (x,y) system relative to the building map.
+Another method is to consider the indoor map as the reference and employ a cartesian coordinate system, x and y, to represent a location.
+Meridian, an indoor location system developed by HP \cite{meridian}, which presents functionalities such as route making and push notifications through its beacons, makes use of its platform for insertion of the indoor maps.
+As such there isn't the outdoor component that google indoor maps has, making it possible to use a (x,y) system relative to the building map.
 
-Although Meridian already provides some extra level of detail in comparison with google indoor maps, there is another example that should be mentioned called OpenStreetMaps.
+Although Meridian already provides some extra level of detail in comparison with google indoor maps, there is another example that should be mentioned, OpenStreetMaps.
 Although it started much like google, providing global data, due to its openness, many indoor projects surfaced such as OpenLevelUP \cite{openlevel}.
-This project makes use of OpenStreetMaps' current indoor tagging scheme \cite{opentagging}, which is intented to describe in the most complete and simple way a building.
-This makes available the number of floors, the type of elemets (room, wall, corridor, etc) and its connectors, like doors and escalators or elevators, allowing to understand clearly the map that is being analysed.
+This project makes use of OpenStreetMaps' current indoor tagging scheme \cite{opentagging}, which is intended to describe in the most complete and simple way a building.
+This makes available the number of floors, the type of elements (room, wall, corridor, etc) and its connectors, like doors and escalators or elevators, allowing to understand clearly the map that is being analysed.
 
@@
 
-In order to understand how a system based on this generic architecture, figure \ref{fig:generic} has to be strutured, it is of relevance to analyse the system component's description and requirements that would allow it to be deployed.
-If one starts by analysing the beacons component, one can describe them as the reference points of the system which are in charge of providing nearby users with information relative to their surroundings.
+In order to understand how a system based on this generic architecture, which can be seen on figure \ref{fig:generic}, has to be structured, it is of relevance to analyse the system component's description and requirements that would allow it to be deployed.
+If one starts by analysing the beacons component, one can describe them as the reference points of the system, which are in charge of providing nearby users with information relative to their surroundings.
 These beacons are always ready and available to the users and upon being contacted, they should inform the smartphone which has initiated the communication of the address of the beacon's associated location server.
@@
 Once a location is requested by the application, the service should communicate with nearby beacons through one of its sensors.
-A service is associated with one and only one technology, as such if the application is to support more than one technlogy, multiple services should be implemented.
-This condition is relevant since the required sensor depends on the chosen technology and because it allows the service to quickly associate the data that it is collecting with the technology, information which must be delivered to the location server.
-Once the beacon data has been collected, the same data is forwarded to the beacon's associated server, together with a tag representing the technology of the beacons.
-Upon obtaining a location, the service should forward the data to the application, which is responsible of later communicating with the map server.
-In terms of requirements, the service needs to have Wi-Fi or at least 3G available for communicating with both the location and maps server, and access to the sensor required to communicate with the nearby beacon, for BLE it would be the bluetooth antenna or for QR it would be the camera.
+A service is associated with one and only one technology, making it so that if the application is to support more than one technology, multiple services should be implemented.
+This condition is relevant since the required sensor is dependent on the chosen technology, and because it allows the service to quickly associate the data that it is collecting with the technology.
+Once the beacon data has been collected, it's forwarded to the beacon's associated server, together with a tag representing the technology of the beacons.
+Upon obtaining a location, the service should pass the data to the application, which is responsible of later communicating with the map server.
+In terms of requirements, the service needs to have Wi-Fi or at least 3G available for communicating with both the location and maps server, as well as access to the sensor required to communicate with the nearby beacon.
+In the case of BLE, the sensor would be the Bluetooth antenna, while for QR it would be the camera.
 
@@
 latitude/longitude or x/y.
-The server should also be able to take input from any of the supported technologies, depends on the beacons in the database, in order to output the user's location.
-From this description one can define the requirements of the location server as having in storage the beacons that are associated it as well as their location, being capable of communicating with the smartphone application and being in accordance with the map server, achieved through providing location description that is the same as that present in the maps server.
+The server needs to be capable of taking input from any of the supported technologies, in order to output the user's location.
+From this description one can define the requirements of the location server as: having in storage the beacons that are associated it as well as their location; being capable of communicating with the smartphone application; and being in accordance with the map server, achieved through providing location description that is the same as that present in the maps server.
 
@@
 
-The solution presented in this paper was made with the objective of creating a generic indoor location system capable of being implemented using any existant indoor system created.
+The solution presented in this paper was made with the objective of creating a generic indoor location system capable of being implemented using any existent indoor system created.
 The generic system's architecture is presented in figure ~\ref{fig:generic} and it's divided in 4 parts: beacon, location server, map server and smartphone application.
@@
 the device's sensors need to be able to capture the data.
-Using the previously presented technologies as examples, these "beacons" could be in fact BLE beacons , Wi-Fi access points, LED lamps or even something as simple as sound, with the data capturing being made through the smartphone's available antennas, camera or microphone.
-The beacon represents a position in the indoor environments and as such it needs to be uniquely identified, be it through its characteristics or associated information.
+Using the previously presented technologies as examples, these "beacons" could be in fact BLE beacons , Wi-Fi access points, LED lamps or even something as simple as sound, with the data collection being made through the smartphone's available antennas, camera or microphone.
+The beacon represents a position in the indoor environments and it is uniquely identified, be it through its characteristics or associated information.
 In addition to these informations, each beacon needs to have an associated location server and provide information about it to the smartphone application.
 
-The location server is an external component where the information relative to the "beacons" utilised is stored.
-As such when a location request arrives from the smartphone, the received data can be translated onto physical locations which will later be used to compute the user's location.
-The algorithm used to obtain such location is entirely dependant on the system's creators.
-Once the location is obtain, it's sent back to the smartphone in addition to the location of the server where it should retrieve the associated map.
+The location server is an external component where the information relative to the beacons is stored.
+As such when a location request arrives from the smartphone, the received data can be translated into physical locations, which will later be used to calculate the user's location.
+Once the location is obtained, it's sent back to the smartphone in addition to the location of the server where it should retrieve the associated map.
 
-The map server represents the architectural block responsible for providing the maps associated to the location of the user, location which was obtained through the location service on the application.
-The used map representation is up to the system creater for as long as it is in accordance with the remaining parts of the system.
-This means that the location provided by the location server needs to be something that is representable on the maps provided and capable of being comprehended by the smartphone application.
+The map server represents the architectural block responsible for providing the maps associated to the location of the user.
+The used map representation is up to the system for as long as it is in accordance with the remaining parts of the system.
+This means that the location provided by the location server needs to be representable on the maps provided and capable of being comprehended by the smartphone application.
 
@@
 It is responsible of discovering and communicating with the beacons existent in its environment through its sensors.
-This communication process allows the smartphone to obtain information of its surrounding and to obtain the address of the location server that is associated to the connected beacon, in order to later forward this same information to it.
-Upon communicating with the location server, the application expects to get back information relative to its positions as well as the address of the map server that is to be contacted.
-Through that address, the application is capable of obtaining the map relative to its positions and display the result to the user.
+This communication process allows the smartphone to obtain information of its surroundings and to obtain the address of the location server that is associated to the connected beacon, in order to later forward this same information to it.
+Upon communicating with the location server, the application expects to get back information relative to its position, as well as the address of the map server that is to be contacted.
+Through that address, the application is capable of obtaining the map relative to its position and display the result to the user.
 
-The presented architecture was structured in a way that it is scalable and it allowed for interoperability.
-The idea behind such an architecture would be let different indoor building use their own version of the system but a user with the associated smartphone application would be able to transit between buildings and use the readily available beacons to always obtain its location.
+The presented architecture was structured in a way that it's scalable and allows for interoperability.
+The idea behind such an architecture would be to let different indoor building implement the system, while a user with the associated smartphone application would be able to transit between buildings and use the readily available beacons to always obtain its location.
 This would be achieved through a common architecture that allowed for self-contained implementations that could be access through a generic smartphone application.
-In order to analyse this assumption it is necessary to look back on the already analysed figure \ref{fig:choices}, which represents the main components of an indoor systems.
-This architecture is capable of externalizing from the smartphone component all the remaining.
-By making the smartphone the central piece of communication, the beacons are required to the ones to provide location data.
-By having a server dedicated to computing the location of an user, the system developers are allowed to use whichever algorithm they wish, since there is no dependency to the application.
-They are also the only ones responsible for having the beacon associated data, for giving those same beacons the location of its server and for computing a location that is in accordance with how they wish to store their maps on the maps server.
+In order to analyse this assumption, it is necessary to take another look at figure \ref{fig:choices}, which displays the main components of an indoor systems.
+This architecture is capable of externalising all the components.
+By making the smartphone the central piece of communication, the beacons are required to be the ones providing location data.
+By having a server dedicated to computing the location of an user, the system is allowed to use any algorithm, since there is no dependency to the application.
+The location server is also the only one responsible for having the beacon associated data.
 The last remaining component is the location description which is completely passed onto the map server and whose only requirement is for it to be in accordance with its associated location server, in terms of location description.
@@
 
-The implementation presented in this paper was created by utilizing the generic indoor location system presented in section \ref{subsec:architecture} and applying it with bluetooth low energy.
-The system's architecture is presented in figure ~\ref{fig:implementation} and is divided three parts: the bluetooth low energy device, in section ~\ref{subsec:beacon} a description of the used technologies and the changes made are present;the server, whose funcionalities and stored information are described in section ~\ref{subsec:server}; and the smartphone application, whose process is described in section ~\ref{subsec:service} alongside figures that show the functional prototype.
-For each of these parts an explanation will be given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
+The implementation presented in this paper was created by employing the generic indoor location system presented in section \ref{subsec:architecture} and applying it with bluetooth low energy.
+The system's architecture is presented in figure \ref{fig:implementation} and is divided three parts: the Bluetooth low energy device, in section \ref{subsec:beacon} a description of the used technologies and the changes made are presented; the server, whose functionalities and stored information are described in section \ref{subsec:server}; and the smartphone application, whose process is described in section \ref{subsec:service} alongside figures that show the functional prototype.
+For each of these parts an explanation is given, containing a description of each of its components specific to the presented system alongside the requirements for each to work.
 
@@
 
-When looking at the BLE's architecture , which can be seen at \ref{fig:BLEarchitecture}, the Generic Access Profile (GAP) is one of the most relevant since it's responsible for working in conjunction with Generic Attribute (GATT) to define the base funcionality of BLE devices.
-The services that are made available by the GAP are: BLE device discovery, connection modes, security, authentication, association models and service discovery.
-In addiction to this, it also defines four different roles to describe a device, allowing for the controllers to be optimized in funtion of the device's desired roles: 
-Broadcaster, role optimized for transmitter-only applications; 
-Observer, role optimized for receiver-only applications and it's complementary to the broadcaster role;
-Peripheral, role optimized for devices that only want to suppot a single connection, allowing for a much less complex controller due to the fact that it only needs to support the slave role and not the master one;
-Central, role supports multiple connections and funtions as the initiator for all of them.
-These connection are all made with Peripheral devices and its controller must support the master role in a connection and allow for more complex funtions, in comparison to the remaining roles.
+For better understanding the beacon-smartphone communication, it is fundamental to comprehend how this technology works.
+All the presented information on BLE was obtained from its core manual \cite{BLECore}.
 
-Another important component is the Attribute Protocol (ATT) Protocol which is responsible for implementing the Peer-to-peer(P2P) protocol between an attribute server and client.
-This communication happens in a dedicated fixed  channel and a server can send through it responses, notifications and indications, while the client can send requests, commands and confirmations.
-The ATT allows the clients to read and write values of attributes on a peer device acting as a server.
-
-The last component is the Generic Attribute (GATT) Profile which is responsible for creating a framework for the ATT, in which it's represented the funcionalities of an ATT server.
-This profile describes the hierarchy of services, characteristics and attributes existent in the server and provides an interface for discovering, reading, writing and indicating service characteristics and profiles.
-
-Bluetooth low energy devices utilize profiles which define the required functionalities of the device.
-It also defines application behaviour and data formats and as such when two devices comply with all the requirements of a Bluetooth profile, application interoperability is achieved.
-Each Bluetooth profiles describes its requirements necessary for devices to create a connection, to find available services and connection information required for making application level connections.
+The smartphone and the beacon are set up as two complementary classes.
+The beacon is defined as a peripheral, a role used to describe devices optimized to support a single connection.
+This role allows it to have a reduced complexity, since it's only required to function as a slave, an entity that is only capable of receiving connections.
+On the other hand, the smartphone is defined by the complementary role, central.
+A central device is capable of supporting multiple connections and is responsible for initiating all of them.
 
@@
 
-The base profile that any Bluetooth system needs to include is the GAP.
-From this point, any additional profile implemented will be a superset of GAP, where GATT is included and specifies the profile hierarchy, or the structure in which profile data is exchanged.
-Figure  ~\ref{fig:profile} shows the hierarchy in a Gatt-based profile, with the profile being the top level and services and characteristics below.
-A profile is composed by one or more services.
-A service is a collection of data and associated behaviors to accomplish a particular function or feature of a device or portions of a device.
+Another important aspect in understanding this technology is the BLE profiles.
+A profile, which can be seen on figure \ref{fig:profile} defines an hierarchy in which a device's available services are organized.
+By doing so, a system's profile ends up defining the applications behaviour and data formats, as well as the manner in which data is exchanged.
+The hierarchy is composed by two parts: Services and Characteristics.
+
+\begin{itemize} 
+\item A profile is composed by one or more services.
+A service is a collection of data and associated behaviours to accomplish a particular function or feature of a device or portions of a device.
+It can be either primary, which provides primary functionalities of a device, or secondary, providing auxiliary functionalities of a device and being referenced from at least one primary service.
 A service is composed of characteristics and/or references to other services.
-A Characteristic is a value that is used in a service that has properties and configuration information that descrive how the value should be accessed as well as information on how to display the value.
-A characteristic is defined by its declaration, its properties, its value and may also be defined by its descriptor, which describes the value or permit configuration of the server relative to the value.
+
+\item A Characteristic is a value that is used in a service that has properties and configuration information that describe how the value should be accessed as well as information on how to display the value.
+A characteristic is defined by its declaration, its properties, its value and may also be defined by its descriptor, which describes the value or permit configuration of
+the server relative to the value. 
+\end{itemize}
 
@@
 
-The beacons that were utilized are Texas Instruments CC2650STK devices which can be visualized in figure ~\ref{fig:beacon}.
-Alongside the device, which comes with a pre-installed bluetooth low energy program capable of giving information on each of its ten sensors through its predefined profiles, there is a texas smartphone application that can connect to a single device and read from its sensors.
-By using the texas Code Composer Studio (CSS), the pre-defined ble profile existant on the device could be altered.
+The used beacons were Texas Instruments CC2650STK devices which can be visualised in figure ~\ref{fig:beacon}.
+Alongside the device, which comes with a pre-installed Bluetooth low energy program capable of giving information on each of its ten sensors through its predefined profiles, there is a texas smartphone application that can connect to a single device and read from its sensors.
+By using the texas Code Composer Studio (CSS), the predefined ble profile existent on the device could be altered.
 Upon further analysis of the profile, a characteristic was found for which the Universal Unique Identifier (UUID) of the service and the characteristic itself was found and as such this was the one that ended up being used to store the device's owner server's address.
-Since the device was already set to work as a pheripheral and it now stored the information relevant to the system, there was no need to do further work.
+Since the device was already set to work as a peripheral and it now stored the information relevant to the system, there was no need to do further work.
 
@@
 The Smartphone application was developed for Android using the Android Studio IDE.
-The Application is divided in two primary functional blocks, the Mock location Provider and the Google Maps Integrated Display, as can be visualized in figure ~\ref{fig:implementation}.
+The Application is divided in two primary functional blocks, the Mock location Provider and the Google Maps Integrated Display, as can be seen in figure ~\ref{fig:implementation}.
 
-The Mock Location Provider is implemenented as if it was a Location provider, such as gps.
+The Mock Location Provider is implemented as if it was a Location provider, such as GPS.
 The application functions works as a listener to a Location provider, in this case it listens to the Mock Provider that was implemented.
 By implementing the whole process of obtaining a location inside a service (the mock provider) , a new level of abstraction is added to the application.
-As such, whenever the application is signaled to obtain the user's current location, a request is made to the associated location provider and the application only need to listen for the answer that eventualy arrives.
+As such, whenever the application is signaled to obtain the user's current location, a request is made to the associated location provider and the application only need to listen for the answer that eventually arrives.
 
@@
 
-The Mock Location Provider incorporates the first three steps present in figure ~\ref{fig:MockProvider}, which will now be explored individualy.
+The Mock Location Provider incorporates the first three steps present in figure ~\ref{fig:MockProvider}, which will now be explored individually.
 The first step indicates the gathering of information of the surroundings of the user's device.
-When a request is made to the provider, a scan for nearby bluetooth low energy devices is made which will put the smartphone in a state of listening for incoming ble advertisement packets.
-During this scanning period, each time a device is found, the advertisement is registered in a list, which have a duplication prevention mechanism implemented.
-Once the period is over, the provider has avaliable a list of all the ble devices within range.
+When a request is made to the provider, a scan for nearby Bluetooth low energy devices is made.
+During this scanning period, each time a device is found, the advertisement is registered in a list, where a duplication prevention mechanism is implemented.
+Once the scanning is complete, the provider has available a list of all the ble devices within range.
 
-The second step involves taking the created list of devices, obtain a server adress and forward the same list to it.
-Once the first step is completed, the provider will analyze each entry at a time.
+The second step involves taking the created list of devices, obtaining a server address and forwarding the same list to it.
+Once the first step is completed, the provider will analyse each entry at a time.
 For each device the provider will attempt to respond to the caught advertisement packet, resulting in a created connection.
@@
 Upon receiving an answer, the list of services is swoop while looking for the service with the wanted UUID.
-If the device doesn't have the UUID that the provider is looking for, it can assume that the paired ble is not a beacon of our system, as such the connection is terminated.
-When the provider identifies that the device has the system's UUID, it requests the device to provide the service's existant characteristics.
+If the device doesn't have the UUID that the provider is looking for, it can assume that the paired ble is not a beacon of the system.
+When the provider identifies that the device has the system's UUID, it requests the device to hand over the service's existent characteristics.
 The provider will receive a list composed of the service's characteristics and it will search in it for the system's characteristics UUID, the one which contains the device's server's address.
-This search has the objective of confirming that the service existant in this device is indeed the one that was implemented for the system and not a device with another service that happened to have the same UUID.
+This search has the objective of confirming that the service present in this device is indeed the one that was implemented for the system and not a service that happened to have the same UUID.
 For any service outside those that are documented in the Bluetooth Special Interest Group (SIG), who have a specific UUID attached to them, the UUID is generated randomly and as such there is a small chance of collision.
 Once the wanted characteristic is found, the provider requests the device to read its value and stores the received value in a list.
-This list will contain the servers of the devices that were found, and for each address there will be a list corresponding to each device , and their corresponding rssi values, from the same owner.
+This list will contain the servers of the devices that were found, and for each address there will be a list corresponding to each device, as well as their corresponding RSSI values.
 In order to quicken the previously described process, the provider keeps in cache the most recent contacted devices.
@@
 The Third step involves a simple client/server tcp interaction.
-The application starts off by formulating the message that it will later on send to the server, this message includes all pairs of device mac address and its associated rssi value captured by the application on the first step.
-Once the message is computed, the application attemps to create a connection with the server at the chosen address at the end of step two.
+The application starts off by creating the message that it will later on send to the server.
+This message includes all pairs of device mac address and its associated rssi value captured by the application on the first step.
+Consequently, the application attempts to create a connection with the server at the chosen address at the end of step two.
 With the connection established, the message is forwarded to the server and the application is put onto blocked state where it awaits for an answer.
-Upon arrival, the answer received is checked for valid location, its information is process and the connection is terminated.
+Upon arrival, the answer received is checked for valid location, its information process and the connection terminated.
 The information contained inside the received message, which was described in section ~\ref{subsec:server}, is then processed into the adequate class capable of storing a geographic location and the same is broadcasted from the mock location provider to its listener.
@@
 The Google Maps Integrated display is implemented using the Google Maps Android API.
-By using Google Maps it was possible to aliviate the weight on application since there wasn't need implement file transfer of indoor building's maps from each dedicated server to each request, which aliviated the servers aswell since there was no need to store its associated building's maps on it.
-Managing the maps was something that was aswell fortunately unnecessary and as such all these features were provided by google maps service.
-By making this development choice, the system as whole became closer to the desired generic approach while making possible for seamless transition between indoor/outdoor maps.
+By using Google Maps it was possible to alleviate the weight on the application since there was no to implement a maps server.
+By making this development choice, the system as a whole became closer to the desired generic approach while making possible for seamless transition between indoor/outdoor maps.
 The only imposed restriction is related to the addiction of new indoor maps onto the google maps, which is possible and well documented but dependent on a third party.
@@
 The Fourth step is called when the application receives a proper location from the request made onto the location provider.
-With the device's location known, a marker is placed on the map with the obtain coordinates (longitude, latitude), the camera is moved in order to be centralized on the position and fully displaying the indoor level map, and the menu visible on figure ~\ref{AppMenu} is updated with the information that is bundled with the received location.
-In order to show the correct level on a multi level building, the "floor" information present in the menu is utilized.
-The API allows for obtaining a list of existant levels on which the maps' camera is focused and as such it's possible to find out to which level the provided locaiton belongs and make so that the application shows it.
+With the device's location known, a marker is placed on the map with the obtain coordinates (longitude, latitude), the camera is moved in order to be centered on the position, while fully displaying the indoor level map.
+The menu visible on figure \ref{fig:AppMenu} is also updated with the information that is bundled with the received location.
+In order to show the correct level on a multi level building, the "floor" information present in the menu is used.
+The API allows for obtaining a list of existent levels on which the maps' camera is focused.
+As such it's possible to find out to which level the provided location belongs, as well as displaying it to the user.
 
@@
 
-The webserver was implement in Python 3 programming language.
+The web server was implement in Python 3 programming language.
 The program implements a simple tcp server capable of receiving multiple request at the same time.
 Each request starts with information sent from an application which include a pair of MAC address and associated RSSI value for each ble device that the same application found.
-Afterwards the list of pairs is filtered in order to remove any existant devices that are not present in the server's database of devices.
+Afterwards the list of pairs is filtered in order to remove any existent devices that are not present in the server's database of devices.
 
@@
 An entry (description of a device) in this database is composed by the device's mac address, its longitude and latitude and its building, floor and room name.
-In addition to the database, a server when initiated can store additional location info such as the server's street, number, zipcode, city and country, allowing this information to be transmited to the client in order to offer an additional level of location description to the user.
+In addition to the database, a server can store additional location info such as the server's street, number, zip-code, city and country, allowing this information to be transmitted to the client, thus offer an additional level of location description to the user.
 The whole location specific information can be visualised in figure ~\ref{fig:AppMenu}.
@@
 Upon having filtered the initial list of pairs, the Cell of Origin (CoO) technique is applied by verifying which of the devices produced a stronger signal on the receptor.
-Upon obtaining the closest device an answer is sent to the application containing all of the information associated to the server and the selected device.
+Upon obtaining the closest device, an answer is sent to the application containing all of the information associated to the server and the selected device.
 
@@
 
-The final state of the implemented system can be visualized in figure ~\ref{fig:AppFocus} and figure ~\ref{fig:AppMenu}.
-The first displays the case of obtaining a location, where the marker has been placed and the camera zoomed
+The final state of the implemented system can be visualised in figure ~\ref{fig:AppFocus} and figure ~\ref{fig:AppMenu}.
+The first displays the case of obtaining a location, where the marker has been placed and the camera zoomed.
 
@@
 
-%% Qualitativamente 
-%% graficos e tal
+In order to analyse the energetic costs of the BLE system implementation a series of tests were planned and executed.
+The tests were conducted with the intention of discovering the costs of the smartphone on standby, the overall cost associated with the system as well as the ones associated to the communication process.
 
-- Energy Consumption in general
+The used smartphone was a Sony Xperia E5, whose battery is rated at 2300 mAh.
+The battery values data collection was made through the android API, which allows one to know the battery percentage of the smartphone.
 
-- Energy consumption vs Cycle time vs Accuracy?
+The tests were all conducted under the same base conditions, the battery was completely charged, without a SIM card and with the screen at maximum brightness.
+They were also conducted over the period of an hour.
 
-- Time for request completion
+The first batch of tests, presented in figure \ref{fig:syson}, display how the smartphone behaves without the system.
+The first test case shows the energetic costs for the smartphone on standby, i.e.
+with all sensors inactive, test \#2 occurs with the Wi-Fi active and without service, while test \#3 was performed with service available.
+The last test case \#4 is useful since it provides the consumption associated to having all the sensors active (Wi-Fi and Bluetooth).
+This value will allow for a better analysis of the system's performance.
 
-- Data Requiremnts?
+\begin{figure} 
+\centering 
+\includegraphics[width=1\linewidth]{figures/system-on.png} 
+\caption[Test case \#1-4: System's energetic costs]{Test case \#1-4: System's energetic costs} 
+\label{fig:syson} 
+\end{figure}
+
+The second batch of tests, presented in figure \ref{fig:syswork}, allowed for the system's operational costs to be analysed.
+As seen in the figure, the parameter that is being tuned is the operation's cycle period, 5,15,30 and 60 seconds, which is tested with one and two devices.
+Through these tests it is possible to visualise the impact of the number of operations on the battery, which is proportional to the number of devices.
+Passing from one to two devices leads to an increase in the order of the 1\% of the total battery.
+This is caused by the increased cost of the BLE communication, since the systems is required to query one extra device each cycle.
+
+\begin{figure} 
+\centering 
+\includegraphics[width=1\linewidth]{figures/system-working.png} 
+\caption[Test case \#5-12: System Working - energetic costs]{Test case \#5-12: System Working - energetic costs} 
+\label{fig:syswork} 
+\end{figure}
+
+In order to better understand the costs associated with the BLE communication component, it is necessary to also understand the costs associated to the network component, i.e.
+communication with both servers.
+Figure \ref{fig:commcosts} allows one to do that, since it shows side-by-side the cost of the whole operation cycle and the costs of the operational cycle without the network communication, i.e.
+the cycle is concluded after concluding the BLE communication.
+The other factor that is tuned is the number of devices which turnout out to be relatively irrelevant since the message's complexity grows slowly with the increase in the number of devices.
+
+\begin{figure} 
+\centering 
+\includegraphics[width=1\linewidth]{figures/system-working.png} 
+\caption[Network Communication's Energetic cost]{Network Communication's Energetic cost} 
+\label{fig:commcosts} 
+\end{figure}
+
+Although the operational costs can seem rather high, the test context isn't implement with any of the possible optimisation techniques.
+When analysing the BLE communication costs, one has to be aware that in each cycle the beacons were contacted, meaning that if the associated address would be kept, only on query would be necessary.
+A small cache-like optimisation would heavily reduce the system's weight, making it so that only new devices would be required to be queried for their owner's server address.
+Regarding the network communication costs, they can be reduced by making use of other sensors to discover when the user is moving, so that only then the positions is searched for, as well as making optimisations regarding the maps component, such as not requesting the same locations multiple times.
 
@@
 The parts that can be changed are the three possibilities previously studied from figure \ref{fig:choices}.
-In case the technlogy is to be change, we will use QR code as an example, the affected areas are the beacons, the smartphone service and the location server.
-By changing the to QR code, the technology changes instantly as there is no more requirement for BLE beacons but instead there will be QR codes placed in fixed places.
+In case the technology is to be change, we will use QR code as an example, the affected areas are the beacons, the smartphone service and the location server.
+By changing to QR codes, the technology changes instantly as there is no more requirement for BLE beacons, but instead there will be QR codes placed in fixed places.
 Each QR code would be required contain information about its associated server.
-The service would change the mechanism utilised to collect data, instead of listening for BLE beacons through its bluetooth antenna, it will open up the camera, so that the user can capture the QR code which would then be forwarded to the location server.
-The location server would have to be prepared for receiving the QR data and obtaining the location relative to that QR code.
+The service would change the mechanism used to collect data, instead of listening for BLE beacons through its Bluetooth antenna, it will open up the camera, so that the user can capture the QR code.
+The location server would have to be prepared to receive the QR data and obtaine the location relative to that QR code.
 
-If the system were to be changed on the location algorithm component, this would only require changes on the location server.
-The algorithms connection has only two connections points with the system, the data that it receives and the location it outputs.
+If the system was to be changed on the location algorithm component, this would only require changes on the location server.
+The algorithm component has only two connections points with the system, the data that it receives and the location it outputs.
 As such, for as long as the new algorithm makes use of the same data and in the end, once it has concluded, it outputs one location of the same format, no other issue should surface.
 
-A modification on the location description utilised would have an impact on both the location and map server.
+A modification on the location description would have an impact on both the location and map server.
 With such a request, the map server is required to change completely the stored maps in order to support the new description method, while the changes on the location server are on the database, since the location of the beacons should change, and on the way the location algorithm processes these same locations.
 
-\section{Future Work}
-\label{sec:future}
+\section{Conclusions} 
+\label{sec:conclusions}
 
-- Conclusions
+In this paper, one has presented GEFILOC, a generic framework architecture for indoor location systems centered around smartphones.
+This architecture intended to present a common ground for the many existing indoor location systems, where the central piece of communication would be the smartphone.
+By making it the central component and using its sensors, one is capable of externalising the different components of a generic indoor system.
+This allows for higher scalability, as well as interoperability between different systems.
+
+An implementation of the architecture was achieved through the usage of the Bluetooth low energy technology and the Android operating system.
+The implemented system was successful, having achieved communication between beacons and smartphone, having implemented a location server capable of host the beacon's data and any location algorithm, and having achieved incorporation of the google maps as its maps provider.
+
+The energetic costs associated to the implemented system were promising although for a complete understanding of its cost in a fully-developed system, a more optimised version would be required.
+Despite this factor, the most relevant cost was the one associated to the network, since it came as a tradeoff for the scalability, which didn't achieve concerning values.
+
+These results indicate that it could be possible to use this architecture in multiple indoor location systems and eventually integrating them into the same smartphone application service.
+This situation would allow users to navigate through buildings, while interacting with whichever system is associated to his vicinity.
 
